---
title: "2-Set (Time Periods) Cluster Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2-Set Clustering:  2000 Cluster & 2017 Cluster

This Rmd replicates the original 2018 work done on two different clusters, 2000 & 2015.

```{r}
# library(extrafont)
# loadfonts(device = "win")

acs_17 <- read.xlsx('Data/acs20175_tr.xlsx')
snc_tracts <- read.xlsx('Data/SNC tracts.xlsx')
housing_snc <- fread('Data/acs20175_new_builds.csv')

snc_tracts_tidy <- snc_tracts %>%
  mutate(GEOID = as.character(GEOID))

housing_snc_tidy <- housing_snc %>%
  mutate(GEOID2 = as.character(GEOID2)) %>%
  select(GEOID2, perc_new_bld)

# Remove extraneous variables
snc_acs_vars <- acs_17 %>%
  transmute(tract = GEOG_UNIT,
            pop_total = POPTOTAL,
            age65up = AGE65UP,
            white_nh = WHITENH,
            median_hhi = MEDIANHHI) %>%
  mutate(perc_poc = (pop_total - white_nh)/pop_total*100,
         perc_65up = age65up/pop_total*100) %>%
  select(-pop_total, -white_nh, -age65up)

snc_17 <- inner_join(snc_acs_vars, snc_tracts_tidy, by = c('tract' = 'GEOID')) # Select SNC tracts

snc_17_acs <- left_join(snc_17, housing_snc_tidy, by = c('tract' = 'GEOID2'))

# Bring in imputed rent
rent_imp <- read_csv('Data/acs_rent_with_imp.csv')

rent <- rent_imp %>%
  mutate(rent = ifelse(is.na(MEDGRENT), rent_imputed, MEDGRENT)) %>%
  select(GEOID, rent) %>%
  mutate(GEOID = as.character(GEOID))

# Bring in median home value
emv <- fread("Data/parcel_emv_2017.csv")

emv_tidy <- emv %>%
  mutate(Tract = as.character(Tract)) %>%
  unique()

tenures <- acs_17 %>%
  select(GEOG_UNIT, OWNEROCC, RENTEROCC) %>%
  rename(tract = GEOG_UNIT,
         owner_occ = OWNEROCC,
         rent_occ = RENTEROCC) %>%
  mutate(rent_tenure = rent_occ/(owner_occ + rent_occ),
         owner_tenure = owner_occ/(owner_occ + rent_occ)) %>%
  select(-rent_occ, -owner_occ) %>%
  unique()

# Join all data (excluding tenures, which don't get standardized) together
snc_rent <- left_join(snc_17_acs, rent, by = c("tract" = "GEOID"))
snc_full <- left_join(snc_rent, emv_tidy, by = c("tract" = "Tract"))

snc_for_cluster <- snc_full %>%
  unique()

# Create z-scores
set.seed(413)

## Results of uncapped scaling
z_no_cap <- as_tibble(scale(snc_for_cluster[-1]))
summary(z_no_cap)

## Results of capped scaling
z_cap <- replace(z_no_cap, z_no_cap > 3, 3)
z_cap <- replace(z_cap , z_cap < -3, -3)
summary(z_cap)

snc_tr_order <- snc_for_cluster %>%
  select(tract)

z_cap_tr <- bind_cols(z_cap, snc_tr_order)

z_cap_tenures <- left_join(z_cap_tr, tenures, by = "tract")

## Create composite of median home value and median rent and drop the intermediaries
z_cap_wt <- z_cap_tenures %>%
  mutate(home_value_rent_comp = (Median_EMV_2017*owner_tenure) + rent*rent_tenure) %>%
  select(-Median_EMV_2017, -owner_tenure, -rent, -rent_tenure)

z_cap_clusts <- z_cap_wt %>%
  select(-tract) # Make sure not to include tract in clustering!

# Create distance matrix of z-scores
z_matrix <- as.matrix(dist(z_cap_clusts), labels = TRUE)
fviz_dist(dist(z_cap_wt)) # Look at distance matrix at a glance

# Label rows and columns with tract name
# colnames(z_matrix) <- snc_tr_order[['tract']]
# rownames(z_matrix) <- snc_tr_order[['tract']]

fviz_nbclust(z_cap_clusts, pam, method = "silhouette")

set.seed(413)
options(scipen=999)
pam_fit_3 <- pam(z_matrix, diss = TRUE, k = 3)
plot(pam_fit_3)

snc_for_viz <- snc_full %>%
  unique() %>%
  select(-tract)

pam_fit_3$data <- z_matrix

require(cluster)
fviz_cluster(pam_fit_3, geom = "point", ellipse.type = "norm") +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  labs(title = "Clusters along Principal Components 1 & 2") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank())

clusters_17 <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_fit_3$clustering))

# clusters_17r <- clusters_17 %>%
#   mutate(color15 = ifelse(cluster==1,"Red",ifelse(cluster==2,"Blue","Gold")))

clusters_and_zdata <- left_join(clusters_17, z_cap_wt, by = "tract")

cluster_viz <- clusters_and_zdata %>%
  rename(`Percent New Builds` = perc_new_bld,
         `Percent People of Color` = perc_poc,
         `Percent 65+` = perc_65up,
         `Median Household Income` = median_hhi,
         `Home Value-Rent Composite` = home_value_rent_comp) %>%
  gather(`Percent New Builds`, `Percent People of Color`, `Percent 65+`, `Median Household Income`, `Home Value-Rent Composite`, key = "variable", value = "value")

cluster_viz %>%
  mutate(cluster = as.factor(cluster)) %>%
ggplot( aes(value, fill = cluster, color = cluster)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  facet_wrap(~variable) +
  labs(title = "Z-score Distributions by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Non-log Z-score Distributions by Cluster.png")

clusters_and_data <- left_join(clusters_17, snc_for_cluster, by = "tract")

```

# Generate plots of non-log-transformed v. log-transformed variables

```{r}
snc_17_long <- snc_full %>%
  gather(perc_poc, perc_new_bld, perc_65up, median_hhi, rent, Median_EMV_2017, key = 'Variable_category', value = 'Variable_n') %>%
  mutate(Variable_log_n = log(Variable_n)) %>%
  gather(Variable_n, Variable_log_n, key = "Transformation", value = "n") %>%
  mutate(Transformation = ifelse(Transformation == "Variable_n", "None", "Logarithmic"))

vars <- list('perc_poc', 'perc_new_bld', 'perc_65up', 'median_hhi', 'rent', 'Median_EMV_2017')

# Generate plotly plots of log-transformed v. no transformation of variables
plotly_gen <- function(filter_var) {

  filter_var <- as.character(filter_var)
  
  plot_subject <- ifelse(filter_var == 'perc_poc', '% POC',
                         ifelse(filter_var  == 'perc_new_bld', '% NEW BUILDS',
                                ifelse(filter_var  == 'perc_65up', '% 65+',
                                       ifelse(filter_var  == 'median_hhi', 'MEDIAN HHI',
                                              ifelse(filter_var == 'rent', "MEDIAN GROSS RENT",
                                                     ifelse(filter_var == 'Median_EMV_2017', "MEDIAN PARCEL EMV", 'OTHER VARIABLE'))))))

  plot <- snc_17_long %>%
  filter(Variable_category == filter_var) %>%
  ggplot(aes(n)) +
  geom_density(alpha = 0.5, fill = '#006666', color = '#006666') +
  labs(title = paste('DISTRIBUTION OF ', plot_subject, ' ACROSS METRO TRACTS'),
       x = plot_subject) +
  theme_minimal() +
  theme(text = element_text(family = "Tw Cen MT", hjust = 0.5),
        plot.title = element_text(hjust = 0.5, size = 20),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = 'white', color = 'white')) +
    facet_wrap(~Transformation)
  
ggplotly(plot)
  
}

purrr::map(vars, plotly_gen)

# See transformation v. non-transformed on their own

plotly_gen_res <- function(filter_var, transformation) {

  filter_var <- as.character(filter_var)
  
  plot_subject <- ifelse(filter_var == 'perc_poc', '% POC',
                         ifelse(filter_var  == 'perc_new_bld', '% NEW BUILDS',
                                ifelse(filter_var  == 'perc_65up', '% 65+',
                                       ifelse(filter_var  == 'median_hhi', 'MEDIAN HHI',
                                              ifelse(filter_var == 'rent', "MEDIAN GROSS RENT",
                                                     ifelse(filter_var == 'Median_EMV_2017', "MEDIAN PARCEL EMV", 'OTHER VARIABLE'))))))

  plot <- snc_17_long %>%
  filter(Variable_category == filter_var & Transformation == transformation) %>%
  ggplot(aes(n)) +
  geom_density(alpha = 0.5, fill = '#006666', color = '#006666') +
  labs(title = paste('DISTRIBUTION OF ', plot_subject, ' ACROSS METRO TRACTS'),
       x = plot_subject) +
  theme_minimal() +
  theme(text = element_text(family = "Tw Cen MT", hjust = 0.5),
        plot.title = element_text(hjust = 0.5, size = 20),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = 'white', color = 'white'))
  
ggplotly(plot)
  
}

plotly_gen_res("perc_poc", "Logarithmic")
plotly_gen_res("perc_poc", "None")

```

# Try clustering with log transformations, same seed (413)

```{r}
snc_log <- snc_for_cluster %>%
  gather(rent, Median_EMV_2017, perc_65up, perc_new_bld, perc_poc, median_hhi, key = "Category", value = "n") %>%
  mutate(n = ifelse(n != 0, log(n), 0)) %>%
  spread(Category, value = n)
  
# Create z-scores
set.seed(413)

## Results of uncapped scaling
z_no_cap_log <- as_tibble(scale(snc_log[-1]))
summary(z_no_cap_log)

## Results of capped scaling
z_cap_log <- replace(z_no_cap_log, z_no_cap_log > 3, 3)
z_cap_log <- replace(z_cap_log , z_cap_log < -3, -3)
summary(z_cap_log)

snc_tr_order <- snc_log %>%
  select(tract)

z_log_tr <- bind_cols(z_cap_log, snc_tr_order)

z_log_tenures <- left_join(z_log_tr, tenures, by = "tract")

## Create composite of median home value and median rent and drop the intermediaries
z_log_wt <- z_log_tenures %>%
  mutate(home_value_rent_comp = (Median_EMV_2017*owner_tenure) + rent*rent_tenure) %>%
  select(-Median_EMV_2017, -owner_tenure, -rent, -rent_tenure)

z_log_clusts <- z_log_wt %>%
  select(-tract) # Make sure not to include tract in clustering!

# Create distance matrix of z-scores
z_log_matrix <- as.matrix(dist(z_log_clusts), labels = TRUE)
fviz_dist(dist(z_log_wt)) # Look at distance matrix at a glance

# Label rows and columns with tract name
# colnames(z_matrix) <- snc_tr_order[['tract']]
# rownames(z_matrix) <- snc_tr_order[['tract']]

fviz_nbclust(z_log_clusts, pam, method = "silhouette")

set.seed(413)
options(scipen=999)
pam_log <- pam(z_log_matrix, diss = TRUE, k = 3)
plot(pam_log)

pam_log$data <- z_log_matrix

medoids <- as_tibble(pam_log$medoids) %>% rename(Medoid = value)
clus_width <- as_tibble(pam_log$silinfo$clus.avg.widths) %>% rename(`Average Width` = value)
pam_log_df <- bind_cols(medoids, clus_width)
av_width <- as_tibble(pam_log$silinfo$avg.width) %>%
  mutate(av_width_equals = "Average cluster width = ",
         value = round(value, 2)) %>%
  unite(label, av_width_equals, value, sep = "")

ggplot(pam_log_df, aes(Medoid, `Average Width`)) +
  geom_bar(stat = "identity", width = 0.2, fill = "#FFCC33") +
  geom_text(aes("207", 0.25), label = av_width$label, family = "Tw Cen MT") +
  labs(title = "Silhouette Width by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
        plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Cluster Silhouettes.png", width = 10, height = 6)

clusters_log <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_log$clustering))
clusters_log_sil <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_log$clustering))

clusters_and_log_z <- left_join(clusters_log, z_log_wt, by = "tract")

cluster_log_viz <- clusters_and_log_z %>%
  rename(`Percent New Builds` = perc_new_bld,
         `Percent People of Color` = perc_poc,
         `Percent 65+` = perc_65up,
         `Median Household Income` = median_hhi,
         `Home Value-Rent Composite` = home_value_rent_comp) %>%
  gather(`Percent New Builds`, `Percent People of Color`, `Percent 65+`, `Median Household Income`, `Home Value-Rent Composite`, key = "variable", value = "value")

cluster_log_viz %>%
  mutate(cluster = as.factor(cluster)) %>%
ggplot( aes(value, fill = cluster, color = cluster)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  facet_wrap(~variable) +
  labs(title = "Z-score Distributions by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Log Z-score Distributions by Cluster.png", width = 10, height = 6)

fviz_cluster(pam_log, geom = "point", ellipse.type = "norm") +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  labs(title = "Clusters along Principal Components 1 & 2") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank())

ggsave("Results/Clusters Along PCs 1 & 2.png", width = 10, height = 6)

```

# Comparing 2017 results to 2015 results

Change Typology | 2000 | 201X
--------------------------------
A | 2 | 2
B | 2 | 3
C | 3 | 2 -or*-
C | 1 | 2
D | 3 | 3
E | 3 | 1 -or*-
E | 2 | 1
F | 1 | 3
G | 1 | 1

*Given that the clusters with most unlike distributions (for most variables) are 2 and 1.

```{r}
x2000 <- read.xlsx("Data/z_scores2000.xlsx")
x2015 <- read.xlsx("Data/z_scores2015.xlsx")

x2000_t <- x2000 %>%
  select(TRACT, cluster) %>%
  rename(cluster_2000 = cluster)

x2015_t <- x2015 %>%
  select(TRACT, cluster) %>%
  rename(cluster_2015 = cluster)

x2017 <- clusters_log %>%
  rename(cluster_2017 = cluster) %>%
  mutate(tract = as.numeric(tract))

change_00_15 <- left_join(x2000_t, x2015_t, by = "TRACT")
change <- left_join(change_00_15, x2017, by = c("TRACT" = "tract"))

typologies <- change %>%
  mutate(Typology_15 = ifelse(cluster_2000 == 2 & cluster_2015 == 2, "A",
                              ifelse(cluster_2000 == 2 & cluster_2015 == 3, "B",
                                     ifelse(cluster_2000 == 3 & cluster_2015 == 2 | cluster_2000 == 1 & cluster_2015 == 2, "C",
                                            ifelse(cluster_2000 == 3 & cluster_2015 == 3, "D",
                                                   ifelse(cluster_2000 == 3 & cluster_2015 == 1 | cluster_2000 == 2 & cluster_2015 == 1, "E",
                                                          ifelse(cluster_2000 == 1 & cluster_2015 == 3, "F",
                                                                 ifelse(cluster_2000 == 1 & cluster_2015 == 1, "G", "other"))))))),
         Typology_17 = ifelse(cluster_2000 == 2 & cluster_2017 == 2, "A",
                              ifelse(cluster_2000 == 2 & cluster_2017 == 3, "B",
                                     ifelse(cluster_2000 == 3 & cluster_2017 == 2 | cluster_2000 == 1 & cluster_2017 == 2, "C",
                                            ifelse(cluster_2000 == 3 & cluster_2017 == 3, "D",
                                                   ifelse(cluster_2000 == 3 & cluster_2017 == 1 | cluster_2000 == 2 & cluster_2017 == 1, "E",
                                                          ifelse(cluster_2000 == 1 & cluster_2017 == 3, "F",
                                                                 ifelse(cluster_2000 == 1 & cluster_2017 == 1, "G", "other"))))))),
         Typology_change = ifelse(Typology_15 != Typology_17, 1, 0))

typologies %>%
  group_by(Typology_change) %>%
  count()

typologies %>%
  filter(is.na(Typology_change))

write_csv(typologies, "Results/Typologies in 2000, 2015, and 2017.csv")

```
