---
title: "Data Development"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import packages

```{r}
library(tidyverse)
library(openxlsx)
library(foreign)
library(data.table)
library(rgdal)
library(sf)
library(viridis)
library(leaflet)
library(mapview)
library(leafpop)
library(bit64)
library(factoextra)
library(cluster)
library(plotly)
library(extrafont)
#font_import() # Only need to do once
#loadfonts(device = 'win')

Sys.setenv(PATH = paste("C:\\Python27\\ArcGIS10.5\\", Sys.getenv("PATH"), sep=";"))
```

# A Note

Parcel data development contains a number of very large files that exceed GitHub limits.  They are contained within the CommDev/Research/Research/Suburban-Neighborhood-Change folder for sake of future reproducibility.

# Percent New Builds

Below is code for pulling the % new builds from Matt's 5-year ACS data file with all variables - CommDev/Research/Research/Census/ACS/Excel (zipped 'all' file); this file is too large to commit to GitHub, so it is not contained here.

```{r}

housing_data <- fread('Data/acs20175_all.csv')

# Built environment variable

housing_snc <- housing_data %>%
  select(B25034e1, B25034e2, B25034e3, B25034e4, SUMLEV, TCFLAG, GEOID, GEOID2, GEONAME) %>%
  mutate(hu_tot_17 = B25034e1, # Total housing units in existence in 2017
            hu_00_17 = B25034e2 + B25034e3 + B25034e4) %>% # Units built 2000-2009, units built 2010-2013, units built 2014 or later
   filter(SUMLEV == 140 & TCFLAG == 1) %>% # Summary level of geography = tract & contained within Twin Cities
   mutate(perc_new_bld = hu_00_17/hu_tot_17*100) %>%
   select(GEOID, GEOID2, GEONAME, perc_new_bld)

#fwrite(housing_snc, 'Data/acs20175_new_builds.csv')

```

# Import data

```{r}
acs_17 <- read.xlsx('Data/acs20175_tr.xlsx')
snc_tracts <- read.xlsx('Data/SNC tracts.xlsx')
housing_snc <- fread('Data/acs20175_new_builds.csv')

snc_tracts_tidy <- snc_tracts %>%
  mutate(GEOID = as.character(GEOID))

housing_snc_tidy <- housing_snc %>%
  mutate(GEOID2 = as.character(GEOID2))

# Remove extraneous variables
snc_vars_17 <- acs_17 %>%
  transmute(tract = GEOG_UNIT,
            med_rent = MEDGRENT)

snc_17 <- inner_join(snc_vars_17, snc_tracts_tidy, by = c('tract' = 'GEOID')) # Select SNC tracts

snc_17_full <- left_join(snc_17, housing_snc_tidy, by = c('tract' = 'GEOID2'))

# Attach to polygon structure
metro_tr <- st_read(dsn = 'Data', layer = 'Census2010TigerTract', quiet = T)
metro_tr_data <- left_join(metro_tr, snc_17_full, by = c('GEOID10' = 'tract'))

# Filter out irrelevant tracts
snc_tracts_geo <- metro_tr_data %>%
  filter(!is.na(perc_poc))

# Revert to polygon
snc_tracts_sp <- as(snc_tracts_geo, 'Spatial')

writeOGR(snc_tracts_sp, "Data", "Tracts_ACS_Data_Prior_to_Rent_Imputation", driver = "ESRI Shapefile")
```

# Visual exploration of data

Quick check to make sure all the variables are there, everything makes sense, and correct tracts are selected.

```{r}
# Use native plotting
plot(snc_tracts_sp)
plot(snc_tracts_geo)

object.size(snc_tracts_sp)

# Create a quick interactive plot
mapview(snc_tracts_geo,
        zcol = 'perc_poc',
        legend = T,
        alpha.regions = 0.5)
```

# Impute tracts with rent tenure but median gross rent values "NA"

Some tracts have rentals, but no estimated median gross rent.  Impute gross rent so it doesn't appear rent is free.  Note this imputation was done after selection of MUSA, non-core tracts.  Ideal imputation would use all available tracts, although this will work for first pass at cluster.

```{python}
import arcpy
from arcpy import env
env.overwriteOutput = True
env.workspace = "N:/CommDev/Research/Research/Suburban-Neighborhood-Change/Project"

# Convert shp to layer
arcpy.MakeFeatureLayer_management("N:/CommDev/Research/Research/Suburban-Neighborhood-Change/Project/Data/Tracts_ACS_Data_Prior_to_Rent_Imputation.shp", "tracts_for_imp_lyr")

arcpy.FillMissingValues_stpm("tracts_for_imp_lyr", "Rent_Interpolated", "med_rnt", "AVERAGE",
                             "K_NEAREST_NEIGHBORS", "", "", "", 8)

```

# Read in and prep 2000 and 2017 data for clustering on the change

```{r}
# Process 2000 data
x2000 <- read_csv("Data/cleaned_variables2000.csv") # Cleaned in previous iteration of project by Dennis Farmer
snc_tr <- read.xlsx("Data/SNC tracts.xlsx")
weight_00 <- read.xlsx("Data/NHGIS00BY10CT.xlsx")

# Get rent tenure (for weighting EMVs for parcels/median gross rent)
tenure_00 <- weight_00 %>% 
  transmute(TRACT = as.numeric(TR10),
            rent_tenure = RENTOCCW/(OWNOCCW+RENTOCCW),
            own_tenure = OWNOCCW/(OWNOCCW+RENTOCCW)) %>% 
  mutate(rent_tenure = ifelse(is.na(rent_tenure), 0, rent_tenure),
         own_tenure = ifelse(is.na(own_tenure), 0, own_tenure))

# Three tracts missing tenure data entirely
tenure_00 %>%
  filter(rent_tenure == 0 | own_tenure == 0)

snc_00 <- inner_join(x2000, snc_tr, by = c("TRACT" = "GEOID"))

snc_00_tidy <- snc_00 %>%
  select(TRACT, percentPOC2000, perSixFive2000, medHHinc2000, MedianValue2000, rent2000, builtSince00_2000) %>%
  unique()


# Process 2017 data

acs_17 <- read.xlsx('Data/acs20175_tr.xlsx')
snc_tracts <- read.xlsx('Data/SNC tracts.xlsx')
housing_snc <- fread('Data/acs20175_new_builds.csv')

snc_tracts_tidy <- snc_tracts %>%
  mutate(GEOID = as.character(GEOID))

housing_snc_tidy <- housing_snc %>%
  mutate(GEOID2 = as.character(GEOID2)) %>%
  select(GEOID2, perc_new_bld)

# Remove extraneous variables
snc_acs_vars <- acs_17 %>%
  transmute(tract = GEOG_UNIT,
            pop_total = POPTOTAL,
            age65up = AGE65UP,
            white_nh = WHITENH,
            median_hhi = MEDIANHHI) %>%
  mutate(perc_poc = (pop_total - white_nh)/pop_total*100,
         perc_65up = age65up/pop_total*100) %>%
  select(-pop_total, -white_nh, -age65up)

snc_17 <- inner_join(snc_acs_vars, snc_tracts_tidy, by = c('tract' = 'GEOID')) # Select SNC tracts

snc_17_acs <- left_join(snc_17, housing_snc_tidy, by = c('tract' = 'GEOID2'))

# Bring in imputed rent
rent_imp <- read_csv('Data/acs_rent_with_imp.csv')

rent <- rent_imp %>%
  mutate(rent = ifelse(is.na(MEDGRENT), rent_imputed, MEDGRENT)) %>%
  select(GEOID, rent) %>%
  mutate(GEOID = as.character(GEOID))

# Bring in median home value
emv <- fread("Data/parcel_emv_2017.csv")

emv_tidy <- emv %>%
  mutate(Tract = as.character(Tract)) %>%
  unique()

tenures <- acs_17 %>%
  select(GEOG_UNIT, OWNEROCC, RENTEROCC) %>%
  rename(tract = GEOG_UNIT,
         owner_occ = OWNEROCC,
         rent_occ = RENTEROCC) %>%
  mutate(rent_tenure = rent_occ/(owner_occ + rent_occ),
         owner_tenure = owner_occ/(owner_occ + rent_occ)) %>%
  select(-rent_occ, -owner_occ) %>%
  unique()

# Join all data (excluding tenures, which don't get standardized) together
snc_17_rent <- left_join(snc_17_acs, rent, by = c("tract" = "GEOID"))
snc_17_full <- left_join(snc_17_rent, emv_tidy, by = c("tract" = "Tract"))

snc_17_tidy <- snc_17_full %>%
  unique() %>%
  rename(emv = Median_EMV_2017) %>%
  mutate(tract = as.numeric(tract))

snc_00_17 <- inner_join(snc_00_tidy, snc_17_tidy, by = c("TRACT" = "tract"))

write_csv(snc_00_17, "Data/2000 and 2017 Values for SNC Tracts and Variables.csv")

```