---
title: "Jolly Tracts"
output: html_notebook
---

------------------------

# Longitudinal Clustering using Jolly tracts for 2000, 2010 and 2017

-------------------------

```{r}
library(tidyverse)
library(kml3d) # for longitudinal clustering
library(stringr) # for num_extract algorithm
library(MASS) # for AIC
library(GGally) # for correlation matrix
library(extrafont) # for fonts for plots
library(pander) # to tidy model output
library(openxlsx)
library(sf) # for geospatial commands
library(nngeo) # for nearest neighbor computation
library(nnet) # for neural nets in multinomial logit
library(data.table)
library(R.utils) # to unzip gz compressed file
library(censusapi) # obtain 2000 data from Census API
```

# Import housing unit data at the block group level; convert to new builds

```{r}
#gunzip('Data/acs20175_all.csv.gz', remove = F) # Unzip compressed gz file, don't remove compressed file
housing_17 <- fread("Data/acs20175_all.csv")

# New builds

new_builds <- housing_17 %>%
  select(B25034e1, B25034e2, B25034e3, B25034e4, B25034e5, SUMLEV, TCFLAG, GEOID, GEOID2, GEONAME) %>%
  mutate(Total_HU_2017 = B25034e1, # Total housing units in existence in 2017
         HU_2010_2017 = B25034e3 + B25034e2,  # Units built 2010 to 2013, units built 2014 or later
         HU_2000_2009 = B25034e4,
         HU_1990_1999 = B25034e5) %>%
   filter(SUMLEV == 150 & TCFLAG == 1) %>% # Summary level of geography = bg & contained within Twin Cities
   mutate(PC_NEW_HU_2000 = HU_1990_1999/(Total_HU_2017-HU_2010_2017-HU_2000_2009),
          PC_NEW_HU_2010 = HU_2000_2009/(Total_HU_2017-HU_2010_2017),
          PC_NEW_HU_2017 = HU_2010_2017/Total_HU_2017) %>%
   select(GEOID, GEOID2, GEONAME, PC_NEW_HU_2000, PC_NEW_HU_2010, PC_NEW_HU_2017)

#fwrite(new_builds, "Data/percent_new_builds_bg.csv")

```

# Take a look at Census API to get 2000 variables

```{r}
apis <- listCensusApis()

# Different ways of finding year 2000 files
apis %>%
  mutate(Year_2000 = str_detect(title, "2000")) %>%
  filter(Year_2000 == TRUE)

apis %>%
  filter(vintage == 2000)

# See which summary file is the file needed
apis %>%
  filter(title == "2000 Decennial: Summary File 1" | title == "2000 Decennial: Summary File 3")

variables_2000_sf1 <- listCensusMetadata(name = "sf1", vintage = 2000, type = "variables")

variables_2000 <- listCensusMetadata(name = "sf3", vintage = 2000, type = "variables")

# Search variables by label
find_variable <- function(search_term) {
variables_2000 %>%
  mutate(Find_term = str_detect(label, search_term)) %>%
  filter(Find_term == TRUE)
}

find_variable("age")

# Search variables by concept
find_concept <- function(search_term) {
  variables_2000 %>%
    mutate(Find_term = str_detect(concept, search_term)) %>%
    filter(Find_term == T)
}

find_concept("Race \\[8")
find_concept("Age")

# Variable crosswalk (by name)
#P006003 - Black
#P006005 - Asian
#P006006 - Native Hawaiian/Pacific Islander
#P006007 - Other race
#P006001 - Total population
#P006002 - White
#P006004 - American Indian/Alaskan Native
find_variable("years")
```

## Import data

```{r}
x2000 <- fread("Data/CEN2000_2010boundaries2.csv")
x2010 <- read.xlsx("Data/Block group data/acs20105_bg.xlsx")
x2017 <- read.xlsx("Data/Block group data/acs20175_bg.xlsx")
jolly_ids <- read_csv("Data/Block group data/jolly_tract_geoids.csv")
jolly_medians <- fread("Data/Block group data/medians_jollytract.csv")
sa <- read.xlsx("Data/Block group data/estimates_V2018_bg.xlsx")

```

## Aggregate 2000 block group data to jolly tracts

```{r}

# Get gross values for each block group - 2000
census_2000 <- x2000 %>%
  dplyr::select(noquote(order(colnames(x2000)))) %>%
  dplyr::select(-GROSSRENT, -HHMEDVAL, -HHMEDINC, -PC_LEPSPAN) %>%
  gather(8:22, key = "Pc_cat", value = "Value_pc") %>% # Collect percentage columns
  mutate(Value_gross = ifelse(Pc_cat == "PC_BACHELORS" | Pc_cat == "PC_GRADPROF", Value_pc * POPOVER25, Value_pc * TOTALPOP),
         Renters = RENT_TENURE * OCC_HOUSING) %>%
  dplyr::select(-Value_pc) %>%
  spread(Pc_cat, value = Value_gross) %>%
  dplyr::select(-AVGHHSIZE, -RENT_TENURE) %>% # remove categories that aren't summed
  mutate(GEOID10 = as.character(GEOID10))

# Add Jolly tract ID for aggregation
jolly_ids_char <- jolly_ids %>% mutate(GEOID_BG = as.character(GEOID_BG))
jolly_00_attr <- left_join(census_2000, jolly_ids_char, by = c("GEOID10" = "GEOID_BG"))

# Aggregate block groups to Jolly tracts by summing - 2000
options(scipen = 999)
jolly_00_sum <- jolly_00_attr %>%
  dplyr::select(-GEOID10, -GEOID_TR) %>%
  filter(!is.na(RID) & RID != -999) %>% # Get rid of block groups not in the study area, including Bethel dorms
  dplyr::select(RID, everything()) %>%
  gather(2:24, key = "Category", value = "Value") %>%
  group_by(RID, Category) %>%
  mutate(Value_summed = sum(Value, na.rm = T)) %>%
  ungroup() %>%
  dplyr::select(-Value) %>%
  unique() %>%
  spread(Category, value = Value_summed)

jolly_00_perc <- jolly_00_sum %>%
  dplyr::select(RID, OCC_HOUSING, TOTALPOP, POPOVER25, HOUSEHOLDS, everything()) %>%
  gather(6:length(jolly_00_sum), key = "Category", value = "Value") %>%
  mutate(Category_prefix = substring(Category, 1, 2)) %>%
  mutate(Percentage_value = ifelse(Category_prefix == "PC" & Category != "PC_BACHELORS" & Category != "PC_GRADPROF", Value/TOTALPOP * 100,
                                   ifelse(Category == "HISPOP", Value/TOTALPOP * 100,
                                          ifelse(Category == "PC_BACHELORS" | Category == "PC_GRADPROF", Value/POPOVER25 *100,
                                                 ifelse(Category == "Renters", Value/OCC_HOUSING * 100,
                                                        ifelse(Category == "HHMOBILE", Value/HOUSEHOLDS *100, Value)))))) %>%
  dplyr::select(RID, Percentage_value, TOTALPOP, Category) %>%
  mutate(Year = 2000) %>%
  unite(Category_yr, Category, Year) %>%
  mutate(Percentage_value = ifelse(RID == -999, 0, Percentage_value)) %>% # Block group containing Bethel dorms is all zeros and returns NAs upon division; 
  spread(Category_yr, value = Percentage_value) %>%
  rename(POPTOTAL_EST_2000 = TOTALPOP,
         HUTOTAL_EST_2000 = HOUSING_UNITS_2000,
         RENT_TENURE_2000 = Renters_2000,
         PC_HISPPOP_2000 = HISPOP_2000,
         PC_HH_NOVEH_2000 = PC_HHNOVEH_2000,
         PC_HH_MOBILE_2000 = HHMOBILE_2000)

# Gross values for 2010
census_2010 <- x2010 %>%
  filter(TCFLAG == 1) %>%
  dplyr::select(GEOID2, POPTOTAL, AGEUNDER18, AGE65UP, WHITENH, BLACKNH, AMINDNH, ASIANNH, PACIFICNH, OTHERNH, MULTRACENH, HISPPOP, NOTHISPPOP, POPOVER25, BACHELORS, GRADPROF, LEP, POV185RATE, RENTEROCC, OWNEROCC, HHMOBILE, HH_NOVEH, HHTOTAL) %>%
  mutate(POP_POV = POV185RATE * POPTOTAL) %>%
  dplyr::select(-POV185RATE)

# Add Jolly RID attribute for aggregation
jolly_10_attr <- left_join(census_2010, jolly_ids_char, by = c("GEOID2" = "GEOID_BG"))

# Aggregate block groups to Jolly tracts by summing
jolly_10_sum <- jolly_10_attr %>%
  filter(!is.na(RID) & RID != -999) %>%
  dplyr::select(-GEOID_TR) %>%
  dplyr::select(GEOID2, RID, everything()) %>%
  gather(3:(length(jolly_10_attr)-1), key = "Census_category", value = "Value") %>%
  group_by(RID, Census_category) %>%
  mutate(Value_summed = sum(Value, na.rm = T)) %>%
  ungroup() %>%
  dplyr::select(-GEOID2, -Value) %>%
  unique() %>%
  spread(Census_category, value = Value_summed)

# Convert sums to percentages where applicable
jolly_10_perc <- jolly_10_sum  %>%
  dplyr::select(RID, POPOVER25, POPTOTAL, HHTOTAL, RENTEROCC, OWNEROCC, NOTHISPPOP, everything()) %>%
  gather(8:length(jolly_10_sum), key = "Census_category", value = "Value") %>%
  mutate(PC = ifelse(Census_category == "BACHELORS" | Census_category == "GRADPROF", Value/POPOVER25 * 100,
                     ifelse(Census_category == "HH_NOVEH" | Census_category == "HHMOBILE", Value/HHTOTAL * 100, Value/POPTOTAL * 100))) %>%
  mutate(RENT_TENURE_2010 = RENTEROCC/(RENTEROCC + OWNEROCC) * 100,
         PC_POC_2010 = (POPTOTAL - NOTHISPPOP)/POPTOTAL * 100) %>%
  dplyr::select(-POPTOTAL, -POPOVER25, -HHTOTAL, -RENTEROCC, -OWNEROCC, -Value, -NOTHISPPOP) %>%
  spread(Census_category, value = PC) %>%
  rename(PC_65UP_2010 = AGE65UP,
         PC_18UNDER_2010 = AGEUNDER18,
         PC_NATIVE_2010 = AMINDNH,
         PC_ASIAN_2010 = ASIANNH,
         PC_BACHELORS_2010 = BACHELORS,
         PC_BLACK_2010 = BLACKNH,
         PC_GRADPROF_2010 = GRADPROF,
         PC_HH_NOVEH_2010 = HH_NOVEH,
         PC_HH_MOBILE_2010 = HHMOBILE,
         PC_HISPPOP_2010 = HISPPOP,
         PC_LEP_2010 = LEP,
         PC_TWORACE_2010 = MULTRACENH,
         PC_OTHERRACE_2010 = OTHERNH,
         PC_HAWAIIANPCFIS_2010 = PACIFICNH,
         PC_POV185_2010 = POP_POV,
         PC_WHITE_2010 = WHITENH
         )

# Gross values for 2017
census_2017 <- x2017 %>%
  filter(TCFLAG == 1) %>%
  dplyr::select(GEOID2, POPTOTAL, AGEUNDER18, AGE65UP, WHITENH, BLACKNH, AMINDNH, ASIANNH, PACIFICNH, OTHERNH, MULTRACENH, HISPPOP, NOTHISPPOP, POPOVER25, BACHELORS, GRADPROF, LEP, POV185RATE, RENTEROCC, OWNEROCC, HHMOBILE, HH_NOVEH, HHTOTAL) %>%
  mutate(POP_POV = POV185RATE * POPTOTAL) %>%
  dplyr::select(-POV185RATE)

# Add Jolly RID attribute for aggregation
jolly_ids_char <- jolly_ids %>% mutate(GEOID_BG = as.character(GEOID_BG))
jolly_17_attr <- left_join(census_2017, jolly_ids_char, by = c("GEOID2" = "GEOID_BG"))

# Aggregate block groups to Jolly tracts by summing
jolly_17_sum <- jolly_17_attr %>%
  filter(!is.na(RID) & RID != -999) %>%
  dplyr::select(-GEOID_TR) %>%
  dplyr::select(GEOID2, RID, everything()) %>%
  gather(3:(length(jolly_10_attr)-1), key = "Census_category", value = "Value") %>%
  group_by(RID, Census_category) %>%
  mutate(Value_summed = sum(Value, na.rm = T)) %>%
  ungroup() %>%
  dplyr::select(-GEOID2, -Value) %>%
  unique() %>%
  spread(Census_category, value = Value_summed)

# Convert sums to percentages where applicable
jolly_17_perc <- jolly_17_sum  %>%
  dplyr::select(RID, POPOVER25, POPTOTAL, HHTOTAL, RENTEROCC, OWNEROCC, NOTHISPPOP, everything()) %>%
  gather(8:length(jolly_10_sum), key = "Census_category", value = "Value") %>%
  mutate(PC = ifelse(Census_category == "BACHELORS" | Census_category == "GRADPROF", Value/POPOVER25 * 100,
                     ifelse(Census_category == "HH_NOVEH" | Census_category == "HHMOBILE", Value/HHTOTAL * 100, Value/POPTOTAL * 100))) %>%
  mutate(RENT_TENURE_2017 = RENTEROCC/(RENTEROCC + OWNEROCC) * 100,
         PC_POC_2017 = (POPTOTAL - NOTHISPPOP)/POPTOTAL * 100) %>%
  dplyr::select(-POPTOTAL, -POPOVER25, -HHTOTAL, -RENTEROCC, -OWNEROCC, -Value, -NOTHISPPOP) %>%
  spread(Census_category, value = PC) %>%
  rename(PC_65UP_2017 = AGE65UP,
         PC_18UNDER_2017 = AGEUNDER18,
         PC_NATIVE_2017 = AMINDNH,
         PC_ASIAN_2017 = ASIANNH,
         PC_BACHELORS_2017 = BACHELORS,
         PC_BLACK_2017 = BLACKNH,
         PC_GRADPROF_2017 = GRADPROF,
         PC_HH_NOVEH_2017 = HH_NOVEH,
         PC_HH_MOBILE_2017 = HHMOBILE,
         PC_HISPPOP_2017 = HISPPOP,
         PC_LEP_2017 = LEP,
         PC_TWORACE_2017 = MULTRACENH,
         PC_OTHERRACE_2017 = OTHERNH,
         PC_HAWAIIANPCFIS_2017 = PACIFICNH,
         PC_POV185_2017 = POP_POV,
         PC_WHITE_2017 = WHITENH
         )

# Get small area pop and housing unit estimates; aggregate to Jolly tracts
sa_jolly <- left_join(sa, jolly_ids_char, by = c("BG10" = "GEOID_BG"))

sa_10_17 <- sa_jolly %>%
  filter(EST_YEAR == 2017 & !is.na(RID) & RID != -999) %>%
  dplyr::select(BG10, RID, EST_YEAR, POPTOTAL_EST, HUTOTAL_EST, POPTOTAL10, HUTOTAL10) %>%
  mutate(POP_NAME = "POPTOTAL_EST",
         HU_NAME = "HUTOTAL_EST",
         EST_YEAR2 = EST_YEAR) %>%
  unite(POP_YR, POP_NAME, EST_YEAR) %>%
  unite(HU_YR, HU_NAME, EST_YEAR2) %>%
  spread(POP_YR, value = POPTOTAL_EST) %>%
  spread(HU_YR, value = HUTOTAL_EST) %>%
  rename(POPTOTAL_EST_2010 = POPTOTAL10,
         HUTOTAL_EST_2010 = HUTOTAL10) %>%
  gather(3:6, key = "Category", value = "Value") %>%
  group_by(RID, Category) %>%
  mutate(Value_summed = sum(Value, na.rm = T)) %>%
  ungroup() %>%
  dplyr::select(-BG10, -Value) %>%
  unique() %>%
  spread(Category, value = Value_summed)

jolly_0010 <- left_join(jolly_00_perc, jolly_10_perc, by = "RID")
jolly_0017 <- left_join(jolly_0010, jolly_17_perc, by = "RID")
jolly_sa <- left_join(jolly_0017, sa_10_17, by = "RID")
jolly_full <- left_join(jolly_sa, jolly_medians, by = c("RID" = "rid")) %>% rename(AA_Jolly_ID = RID)

cluster_3 <- jolly_full %>%
  dplyr::select(noquote(order(colnames(jolly_full)))) %>%
  rename(Jolly_ID = AA_Jolly_ID)

# fwrite(cluster_3, "Data/Block group data/Jolly tract 3-Timepoint Cluster Variables.csv")
```

# Impute missing hhi's, emv's, and rents

```{r}
# See where NAs lie
summary(jolly_medians) 

# Gross rent
summary(lm(GROSSRENT_2017 ~ RENT_TENURE_2017 + HHMEDINC_2017 + HUTOTAL_EST_2017 + POPTOTAL_EST_2017 + PC_POC_2017 + PC_LEP_2017 + PC_HH_MOBILE_2017, data = cluster_3))

# Home value
summary(lm(HHMEDVAL_2000 ~ HHMEDINC_2000, data = cluster_3))
summary(lm(HHMEDVAL_2010 ~ HHMEDINC_2010, data = cluster_3))
summary(lm(HHMEDVAL_2017 ~ HHMEDINC_2017, data = cluster_3))



ggplot(cluster_3, aes(HHMEDVAL_2000, HHMEDINC_2000)) +
  geom_point()
ggplot(cluster_3, aes(HHMEDVAL_2010, HHMEDINC_2010)) +
  geom_point()
ggplot(cluster_3, aes(HHMEDVAL_2017, HHMEDINC_2017)) +
  geom_point()
```

# Compare relationships between Census variables across timepoints

```{r}
x00_17 <- read_csv("Data/25-Variable 2000, 2010-2017 Clustering Dataset.csv")

# Relationship between hhi and emv, Census tracts
summary(lm(MEDHOMEVAL_2000 ~ MEDIANHHI_2000, data = x00_17)) #39%
summary(lm(MEDHOMEVAL_2010 ~ MEDIANHHI_2010, data = x00_17)) #53%
summary(lm(MEDHOMEVAL_2017 ~ MEDIANHHI_2017, data = x00_17)) #57%

# Relationship between medgrent and tenure, Census tracts
summary(lm(MEDGRENT_2000 ~ RENT_TENURE_2000 + MEDIANHHI_2000, data = x00_17)) #27%
summary(lm(MEDGRENT_2010 ~ RENT_TENURE_2010 + MEDIANHHI_2010, data = x00_17)) #36%
summary(lm(MEDGRENT_2017 ~ RENT_TENURE_2017 + MEDIANHHI_2017, data = x00_17)) #45%
```

# Correlation matrix for Jolly tracts at 2000 

```{r}
ggplot <- function(...)
ggplot2::ggplot(...) + scale_color_brewer(palette="Purples") + scale_fill_brewer(palette="Purples")
unlockBinding("ggplot",parent.env(asNamespace("GGally")))
assign("ggplot",ggplot,parent.env(asNamespace("GGally")))

graph_corr <- ggpairs(jolly_2000) #, mapping = aes(color = clusters)),
                      # columns = c('percentPOC2000', 'perc_poc', 'perSixFive2000', 'perc_65up', 'medHHinc2000', 'median_hhi', 'MedianValue2000', 'emv', 'builtSince00_2000', 'perc_new_bld', "clusters"),
                      # columnLabels = c('% POC 00', '% POC 17', '% 65+ 00', '% 65+ 17', 'HHI 00', 'HHI 17', 'EMV 00', 'EMV 17', 'New Builds 00', 'New Builds 17', "clusters"))

graph_corr <- graph_corr + theme(text = element_text(family = "Tw Cen MT"),
                                 strip.background = element_blank(),
                                 panel.background = element_rect(fill = "white"),
                                 panel.grid = element_line(color = "white"))

graph_corr

ggsave("Results/Longitudinal Clustering/Jolly Tracts/Correlation Matrix for Jolly Tracts 2000.png", width = 20, height = 12)
```

# Compare correlations at 2000 timepoint, Jolly tracts v. Census

```{r}
x00_17 <- read_csv("Data/25-Variable 2000, 2010-2017 Clustering Dataset.csv")

census_2000 <- x00_17 %>%
  gather(2:length(x00_17), key = "Category", value = "Value") %>%
  separate(Category, into = c("Cat", "Year"), sep = -4) %>%
  filter(Year == 2000) %>%
  unite(Category, Cat, Year, sep = "") %>%
  mutate(Value = as.numeric(Value)) %>%
  group_by(Category) %>%
  mutate(Value = ifelse(is.na(Value), median(Value, na.rm = T), Value)) %>% # Quick imputation
  ungroup() %>%
  spread(Category, value = Value) %>%
  dplyr::select(-Tract) %>%
  rename(GROSSRENT_2000 = MEDGRENT_2000,
         HHMEDINC_2000 = MEDIANHHI_2000,
         HHMEDVAL_2000 = MEDHOMEVAL_2000)

# Correlation Matrices for 2000 jolly tract data
jolly_2000 <- cluster_3 %>%
  gather(2:length(cluster_3), key = "Category", value = "Value") %>%
  separate(Category, into = c("Cat", "Year"), sep = -4) %>%
  filter(Year == 2000) %>%
  unite(Category, Cat, Year, sep = "") %>%
  mutate(Value = as.numeric(Value)) %>%
  spread(Category, value = Value) %>%
  dplyr::select(-Jolly_ID) %>%
  mutate(GROSSRENT_2000 = as.numeric(GROSSRENT_2000)) %>%
  mutate(GROSSRENT_2000 = ifelse(is.na(GROSSRENT_2000), median(GROSSRENT_2000, na.rm = T), GROSSRENT_2000)) #quick imputation

cor_jolly_2000 <- cor(jolly_2000) %>% as.data.frame()
cor_jolly_2000$VariableA <- rownames(cor_jolly_2000)

cor_jolly_2000_full <- cor_jolly_2000 %>%
  dplyr::select(VariableA, everything()) %>%
  gather(2:length(cor_jolly_2000), key = "VariableB", value = "Correlation_Jolly") %>%
  filter(VariableA != VariableB)

cor_census_2000 <- cor(census_2000) %>% as.data.frame()
cor_census_2000$VariableA <- rownames(cor_census_2000)

cor_census_2000_full <- cor_census_2000 %>%
  dplyr::select(VariableA, everything()) %>%
  gather(2:length(cor_census_2000), key = "VariableB", value = "Correlation_Census") %>%
  filter(VariableA != VariableB)

corr_compare <- full_join(cor_jolly_2000_full, cor_census_2000_full, by = c("VariableA", "VariableB"))

corr_compare_tidy <- corr_compare %>%
  unite(Variables, VariableA, VariableB, sep = "::") %>%
  gather(Correlation_Jolly, Correlation_Census, key = "Tract_type", value = "Correlation")

corr_compare_diff <- corr_compare %>%
  mutate(Difference = Correlation_Census-Correlation_Jolly)

write_csv(corr_compare_tidy, "Results/Longitudinal Clustering/Jolly Tracts/Correlation Comparisons between Jolly Tracts and Census Tracts in Year 2000.csv")
write_csv(corr_compare_diff, "Results/Longitudinal Clustering/Jolly Tracts/Correlation Differences between Census and Jolly Tracts in Year 2000.csv")
```

# Compare correlations at 2017 timepoint, Jolly tracts v. Census

```{r}
#x00_17 <- read_csv("Data/25-Variable 2000, 2010-2017 Clustering Dataset.csv")

census_2017 <- x00_17 %>%
  gather(2:length(x00_17), key = "Category", value = "Value") %>%
  separate(Category, into = c("Cat", "Year"), sep = -4) %>%
  filter(Year == 2017) %>%
  unite(Category, Cat, Year, sep = "") %>%
  mutate(Value = as.numeric(Value)) %>%
  group_by(Category) %>%
  mutate(Value = ifelse(is.na(Value), median(Value, na.rm = T), Value)) %>% # Quick imputation
  ungroup() %>%
  spread(Category, value = Value) %>%
  dplyr::select(-Tract) %>%
  rename(GROSSRENT_2017 = MEDGRENT_2017,
         HHMEDINC_2017 = MEDIANHHI_2017,
         HHMEDVAL_2017 = MEDHOMEVAL_2017)

# Correlation Matrices for 2000 jolly tract data
jolly_2017 <- cluster_3 %>%
  gather(2:length(cluster_3), key = "Category", value = "Value") %>%
  separate(Category, into = c("Cat", "Year"), sep = -4) %>%
  filter(Year == 2017) %>%
  unite(Category, Cat, Year, sep = "") %>%
  mutate(Value = as.numeric(Value)) %>%
  spread(Category, value = Value) %>%
  dplyr::select(-Jolly_ID) %>%
  mutate(GROSSRENT_2017 = as.numeric(GROSSRENT_2017)) %>%
  mutate(GROSSRENT_2017 = ifelse(is.na(GROSSRENT_2017), median(GROSSRENT_2017, na.rm = T), GROSSRENT_2017)) #quick imputation

cor_jolly_2017 <- cor(jolly_2017) %>% as.data.frame()
cor_jolly_2017$VariableA <- rownames(cor_jolly_2017)

cor_jolly_2017_full <- cor_jolly_2017 %>%
  dplyr::select(VariableA, everything()) %>%
  gather(2:length(cor_jolly_2017), key = "VariableB", value = "Correlation_Jolly") %>%
  filter(VariableA != VariableB)

cor_census_2017 <- cor(census_2017) %>% as.data.frame()
cor_census_2017$VariableA <- rownames(cor_census_2017)

cor_census_2017_full <- cor_census_2017 %>%
  dplyr::select(VariableA, everything()) %>%
  gather(2:length(cor_census_2000), key = "VariableB", value = "Correlation_Census") %>%
  filter(VariableA != VariableB)

corr_compare <- full_join(cor_jolly_2017_full, cor_census_2017_full, by = c("VariableA", "VariableB"))

corr_compare_tidy <- corr_compare %>%
  unite(Variables, VariableA, VariableB, sep = "::") %>%
  gather(Correlation_Jolly, Correlation_Census, key = "Tract_type", value = "Correlation")

corr_compare_diff <- corr_compare %>%
  mutate(Difference = Correlation_Census-Correlation_Jolly)

# write_csv(corr_compare_tidy, "Results/Longitudinal Clustering/Jolly Tracts/Correlation Comparisons between Jolly Tracts and Census Tracts in Year 2017.csv")
# write_csv(corr_compare_diff, "Results/Longitudinal Clustering/Jolly Tracts/Correlation Differences between Census and Jolly Tracts in Year 2017.csv")
```

# Get order of variables and assignment for clustering algorithm

```{r}
# cluster_3 <- fread("Data/Block group data/Jolly tract 3-Timepoint Cluster Variables.csv")

# Get order for clustering
assign <- cluster_3 %>%
  gather(everything(), key = "Variable_name", value = "var_value") %>%
  dplyr::select(Variable_name) %>%
  unique() %>%
  mutate(ID = row_number()) %>%
  filter(Variable_name != "Jolly_ID") %>%
  separate(Variable_name, into = c("Variable", "Year"), sep = -4) %>%
  mutate(Year = as.numeric(Year)) %>%
  group_by(Variable) %>%
  mutate(Year_min = min(Year),
         Year_max = max(Year)) %>%
  ungroup() %>%
  filter(Year == Year_min | Year == Year_max) %>%
  dplyr::select(Variable, ID) %>%
  group_by(Variable) %>%
  summarize(Assignments = paste(ID, collapse = "_")) %>%
  separate(Assignments, into = c("Min", "Max"), sep = "_") %>%
  mutate(Min = as.numeric(Min),
         Max = as.numeric(Max)) %>%
  ungroup() %>%
  filter(Variable != "PC_WHITE_20") # Don't include % white in clustering

# Create list of cluster variables for clustering
time_in_data <- vector("list", nrow(assign))
for_length <- 1:nrow(assign)

for (i in for_length) {
  
  assign_slice <- assign %>% slice(i)
  time_in_data[[i]] <- assign_slice$Min:assign_slice$Max

}

names(time_in_data) <- assign$Variable

```

# Longitudinal clustering

```{r}
cld_snc_3 <- cld3d(cluster_3, timeInData = time_in_data)

set.seed(410)
kml3d(cld_snc_3, nbClusters = 8) # Any number of clusters exceeding 6 must be specified using nbClusters
#choice(cld_snc)

# Take a look at 8 clusters
set.seed(410)
cluster_3$clusters <- getClusters(cld_snc_3, 8)

cluster_3_snc <- cluster_3 %>%
  filter(!is.na(clusters)) %>%
  dplyr::select(Jolly_ID, clusters, everything()) %>%
  gather(3:length(cluster_3), key = "Variable_year", value = "Value") %>%
  mutate(Value = as.numeric(Value)) %>%
  separate(Variable_year, into = c("Variable", "Year"), sep = -4) %>%
  separate(Variable, into = c("Variable", "underscore"), sep = -1) %>%
  dplyr::select(-underscore)

fwrite(cluster_3_snc, "Results/Longitudinal Clustering/Jolly Tracts/8-Clusters 3-Timepoints.csv")
```

# Create Jolly tract shapefile

```{r}
metro_bg <- st_read("Data/Block group data/Census2010TigerBlockGroup.shp")

metro_jolly <- left_join(metro_bg, jolly_ids_char, by = c("GEOID10" = "GEOID_BG"))

# Run boundary dissolve within each Jolly tract and view
library(mapview)
metro_jolly %>%  
  split(.$RID) %>% 
  lapply(st_union) %>% 
  do.call(c, .) %>% # bind the list element to a single sfc
  st_cast() %>% # mapview doesn't like GEOMETRY -> cast to MULTIPOLYGON
  mapview()

jollies <- metro_jolly %>%
  split(.$RID) %>%
  lapply(st_union) %>%
  do.call(c, .) 

#st_write(jollies, "Data/Block group data/Jolly tracts.shp")

```

