---
title: "2010 Median EMV by Tract (+ 2000 & 2017 EMVs)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(tidyverse)
library(odbc)
library(DBI)
library(dbplyr)
library(dplyr)
library(sf)
```

----------------------------

# 2010 EMVs

----------------------------

# Connect to GIS SDE; pull 2011 and 2010 parcel datasets

```{r}
# the simplest, goes through master - guest
con <- dbConnect(odbc::odbc(), "GIS")
odbc::dbListTables(con, catalog_name = "GISLibrary", database = "GISLibrary", dbms.name = "dbo")
# odbcListObjects(con, catalog = "GISLibrary", schema = "dbo")

# Get 7-county parcels from 2011
parcels_11 <- dbGetQuery(con, "SELECT * FROM GISLibrary.dbo.Parcels2011All7Points")
parcels_10 <- dbGetQuery(con, "SELECT * FROM GISLibrary.dbo.Parcels2010All7Points")

```

```{r}
# Look at variables in dataset
parcels_11 %>%
  head()

# Check year builds included - includes 2011 + 2012, so will need to remove
levels(as.factor(parcels_11$YEAR_BUILT))

# Look at unique land use descriptions
levels(as.factor(parcels_11$USE1_DESC))

# Look at unique land use descriptions from Dennis' 2015 dataset
emv_15 <- fread("Data/Parcel/EMV 2015 Parcel Work/Final_Res_ParcelPts_2015.txt")

luc <- emv_15 %>% 
  dplyr::select(USE1_DESC) %>%
  unique() %>%
  mutate()

class(emv_10$YEAR_BUILT)
class(parcels_10$COUNTY_ID)
levels(as.factor(parcels_11$COUNTY_ID))

emv_15 %>%
  filter(COUNTY_ID == "123") %>%
  select(USE1_DESC) %>%
  unique() %>%
  filter(USE1_DESC == "Res 1 unit")
```

# 2010 EMVs

```{r}
cols <- c("PIN", "EMV_TOTAL", "COUNTY_ID")

# EMVs for Anoka, Hennepin - according to metadata, these are the 2010 EMVs, taxes payable in 2011; use land use descriptions from Dennis' dataset; eliminate builds after 2010; keep only homesteaded (owner-occupied)
emv_10_ahr <- data.table(parcels_11)[COUNTY_ID == "053" | COUNTY_ID == "123" | COUNTY_ID == "003"][USE1_DESC %in% luc$USE1_DESC][YEAR_BUILT < 2011][, ..cols]

# EMVs for Carver, Dakota, Scott and Washington - according to metadata, these are the 2010 EMVs (except Scott - 2009 EMV, taxes payable 2010)
emv_10_cdsw <- data.table(parcels_10)[COUNTY_ID == "019" | COUNTY_ID == "037" | COUNTY_ID == "139" | COUNTY_ID == "163"][USE1_DESC %in% luc$USE1_DESC][YEAR_BUILT < 2011][, ..cols] # Carver didn't fill in the Homestead status column in 2010

# EMVs for Ramsey (Ramsey's land use descriptions changed between 2010 and 2015)
ramsey <- data.table(parcels_11)[COUNTY_ID == "123"][HOMESTEAD == "Y"][, ..cols]

emv_10 <- rbind(emv_10_ahr, emv_10_cdsw, ramsey)

# fwrite(emv_10, "Data/Parcel/EMV 2010 Parcel Data/2010_SFD_PINs.csv")
```

# Import parcel data as txt files with tract attribute; select only homesteaded SFD PINs

```{r}
# Import parcel data with tract attribute
setwd("EMV 2010 Parcel Data")
csv_files <- list.files(pattern = "*.txt")

parcels <- lapply(csv_files, function(file){
  
  fread(file, select = c("PIN", "GEOID10"))
  
        })

parcels_dt <- do.call("rbind", parcels)
```

# Aggregate to 2010 medians

```{r}
# Import 2010 SFD PINs
# emv_10 <- fread("EMV 2010 Parcel Data/2010_SFD_PINs.csv")

parcels_tract <- left_join(pins_2010, parcels_dt, by = "PIN")

# Left join on pins_2010 (in brackets)
setkey(parcels_dt, "PIN")
setkey(emv_10, "PIN")

parcels_tract <- parcels_dt[emv_10]

median_emv <- parcels_tract[, EMV_2010 := median(EMV_TOTAL),
                            by = GEOID10][, GEOID10, EMV_2010] %>% unique()

# fwrite(median_emv, "EMV 2010 Parcel Data/Median 2010 EMV by Tract.csv")
```

----------------------------

# 2017 EMVs

----------------------------

This procedure fails to pull parcels for a multitude of Hennepin tracts.  Use Dennis' PINs from 2015 instead, and impute missing values.

# Connect to GIS SDE; pull 2018 and 2017 parcel datasets
# Bring in Dennis' 2015 dataset for land use descriptions

```{r}
# the simplest, goes through master - guest
con <- dbConnect(odbc::odbc(), "GIS")
odbc::dbListTables(con, catalog_name = "GISLibrary", database = "GISLibrary", dbms.name = "dbo")
# odbcListObjects(con, catalog = "GISLibrary", schema = "dbo")

# Get 7-county parcels from 2011
hennepin_18 <- dbGetQuery(con, "SELECT * FROM GISLibrary.dbo.Parcels2018HennepinPoints")
ramsey_18 <- dbGetQuery(con, "SELECT * FROM GISLibrary.dbo.Parcels2018RamseyPoints")
parcels_17 <- dbGetQuery(con, "SELECT * FROM GISLibrary.dbo.Parcels2017All7Points")

# Look at unique land use descriptions from Dennis' 2015 dataset
emv_15 <- fread("Data/Parcel/EMV 2015 Parcel Work/Final_Res_ParcelPts_2015.txt")

luc <- emv_15 %>% 
  dplyr::select(USE1_DESC) %>%
  unique() %>%
  mutate()

```

```{r}
# Look at variables in dataset
parcels_11 %>%
  head()

# Check year builds included - includes 2011 + 2012, so will need to remove
levels(as.factor(parcels_11$YEAR_BUILT))

# Look at unique land use descriptions
levels(as.factor(parcels_11$USE1_DESC))

class(emv_10$YEAR_BUILT)
class(parcels_10$COUNTY_ID)
levels(as.factor(parcels_11$COUNTY_ID))

emv_15 %>%
  filter(COUNTY_ID == "123") %>%
  select(USE1_DESC) %>%
  unique() %>%
  filter(USE1_DESC == "Res 1 unit")
```

# 2017 EMVs

```{r}
cols <- c("PIN", "EMV_TOTAL")

# EMVs for Anoka, Carver, Dakota, Scott and Washington - according to metadata, these are the 2017 EMVs; keep only homesteaded (owner-occupied); use land use descriptions from Dennis' dataset; eliminate builds after 2018
emv_17_adsw <- data.table(parcels_17)[COUNTY_ID == "003" | COUNTY_ID == "037" | COUNTY_ID == "139" | COUNTY_ID == "163"][HOMESTEAD == "Y"][USE1_DESC %in% luc$USE1_DESC][YEAR_BUILT < 2018][, ..cols]

carver_17 <- data.table(parcels_17)[COUNTY_ID == "019"][USE1_DESC %in% luc$USE1_DESC][YEAR_BUILT < 2018][, ..cols]

# EMVs for Ramsey
ramsey_17 <- data.table(ramsey_18)[HOMESTEAD == "Yes"][USECLASS1 %in% luc$USE1_DESC][YEAR_BUILT < 2018][, ..cols]

# EMVs for Hennepin
hennepin_17 <- data.table(hennepin_18)[HOMESTEAD == "Yes"][USECLASS1 %in% luc$USE1_DESC][YEAR_BUILT < 2018][, ..cols]

emv_17 <- rbind(emv_17_adsw, carver_17, ramsey_17, hennepin_17)

# fwrite(emv_17, "Data/Parcel/EMV 2017 Parcel Data/2017_SFD_PINs.csv")

```

# Import parcel data as txt files with tract attribute; select only homesteaded SFD PINs

```{r}
# Import parcel data with tract attribute
setwd("Data/Parcel/EMV 2017 Parcel Data/Parcels with Tracts")
csv_files <- list.files(pattern = "*.csv")

parcels <- lapply(csv_files, function(file){
  
  fread(file, select = c("PIN", "GEOID10"))
  
        })

parcels_dt <- do.call("rbind", parcels)
```

# Aggregate to 2017 medians

```{r}
# Import 2010 SFD PINs
# emv_17 <- fread("Data/EMV 2017 Parcel Data/2017_SFD_PINs.csv")

parcels_tract <- left_join(emv_17, parcels_dt, by = "PIN")

# Left join on pins_2017 (in brackets)
setkey(parcels_dt, "PIN")
setkey(emv_17, "PIN")

parcels_tract <- parcels_dt[emv_17]

median_emv <- parcels_tract[, EMV_2017 := median(EMV_TOTAL),
                            by = GEOID10][, GEOID10, EMV_2017] %>% unique()

fwrite(median_emv, "Data/Parcel/EMV 2017 Parcel Data/Median 2017 EMV by Tract.csv")
```

# Collect 2000, 2010, and 2017 medians together (inflation-adjusted in *Longitudinal Clustering.Rmd*)

```{r}
emv_10 <- fread("Data/Parcel/EMV 2010 Parcel Data/Median 2010 EMV by Tract.csv")
emv_17 <- fread("Data/Parcel/EMV 2017 Parcel Data/Median 2017 EMV by Tracts.csv")
emv_00 <- fread("Data/cleaned_variables2000.csv")

emv_00_tidy <- emv_00 %>%
  select(MedianValue2000, TRACT) %>%
  rename(Tract = TRACT,
         EMV_2000 = MedianValue2000)

emv_00_10 <- left_join(emv_00_tidy, emv_10, by = c("Tract" = "GEOID10"))
emv_00_17 <- left_join(emv_00_10, emv_17, by = "Tract")

# fwrite(emv_00_17, "N:\\CommDev\\Research\\Research\\Suburban-Neighborhood-Change\\Project\\Data\\Tract-level\\emv_00_10_17.csv")
```
