---
title: "SNC Update - 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Import packages

```{r}
library(tidyverse)
library(openxlsx)
library(foreign)
library(data.table)
library(rgdal)
library(sf)
library(viridis)
library(leaflet)
library(mapview)
library(leafpop)
library(bit64)
library(factoextra)
library(cluster)
library(plotly)
library(extrafont)
#font_import() # Only need to do once
#loadfonts(device = 'win')

Sys.setenv(PATH = paste("C:\\Python27\\ArcGIS10.5\\", Sys.getenv("PATH"), sep=";"))
```

```{r}
# Below is code for pulling the % new builds from Matt's 5-year ACS data file with all variables - CommDev/Research/Research/Census/ACS/Excel (zipped 'all' file)

# housing_data <- fread('Data/acs20175_all.csv')
# 
# # Built environment variable
# 
# housing_snc <- housing_data %>%
#   select(B25034e1, B25034e2, B25034e3, B25034e4, SUMLEV, TCFLAG, GEOID, GEOID2, GEONAME) %>%
#   mutate(hu_tot_17 = B25034e1, # Total housing units in existence in 2017
#             hu_00_17 = B25034e2 + B25034e3 + B25034e4) %>% # Units built 2000-2009, units built 2010-2013, units built 2014 or later
#   filter(SUMLEV == 140 & TCFLAG == 1) %>% # Summary level of geography = tract & contained within Twin Cities
#   mutate(perc_new_bld = hu_00_17/hu_tot_17*100) %>%
#   select(GEOID, GEOID2, GEONAME, perc_new_bld)

# fwrite(housing_snc, 'Data/acs20175_new_builds.csv')

```

# Import data

```{r}
acs_17 <- read.xlsx('Data/acs20175_tr.xlsx')
snc_tracts <- read.xlsx('Data/SNC tracts.xlsx')
housing_snc <- fread('Data/acs20175_new_builds.csv')

snc_tracts_tidy <- snc_tracts %>%
  mutate(GEOID = as.character(GEOID))

housing_snc_tidy <- housing_snc %>%
  mutate(GEOID2 = as.character(GEOID2))

# Remove extraneous variables
snc_vars_17 <- acs_17 %>%
  transmute(tract = GEOG_UNIT,
            med_rent = MEDGRENT)

snc_17 <- inner_join(snc_vars_17, snc_tracts_tidy, by = c('tract' = 'GEOID')) # Select SNC tracts

snc_17_full <- left_join(snc_17, housing_snc_tidy, by = c('tract' = 'GEOID2'))

# Attach to polygon structure
metro_tr <- st_read(dsn = 'Data', layer = 'Census2010TigerTract', quiet = T)
metro_tr_data <- left_join(metro_tr, snc_17_full, by = c('GEOID10' = 'tract'))

# Filter out irrelevant tracts
snc_tracts_geo <- metro_tr_data %>%
  filter(!is.na(perc_poc))

# Revert to polygon
snc_tracts_sp <- as(snc_tracts_geo, 'Spatial')

writeOGR(snc_tracts_sp, "Data", "Tracts_ACS_Data_Prior_to_Rent_Imputation", driver = "ESRI Shapefile")
```

# Visual exploration of data

Quick check to make sure all the variables are there, everything makes sense, and correct tracts are selected.

```{r}
# Use native plotting
plot(snc_tracts_sp)
plot(snc_tracts_geo)

object.size(snc_tracts_sp)

# Create a quick interactive plot
mapview(snc_tracts_geo,
        zcol = 'perc_poc',
        legend = T,
        alpha.regions = 0.5)
```

# Impute tracts with rent tenure but median gross rent values "NA"

Some tracts have rentals, but no estimated median gross rent.  Impute gross rent so it doesn't appear rent is free.  Note this imputation was done after selection of MUSA, non-core tracts.  Ideal imputation would use all available tract, although this will work for first pass at cluster.

```{python}
import arcpy
from arcpy import env
env.overwriteOutput = True
env.workspace = "N:/CommDev/Research/Research/Suburban-Neighborhood-Change/Project"

# Convert shp to layer
arcpy.MakeFeatureLayer_management("N:/CommDev/Research/Research/Suburban-Neighborhood-Change/Project/Data/Tracts_ACS_Data_Prior_to_Rent_Imputation.shp", "tracts_for_imp_lyr")

arcpy.FillMissingValues_stpm("tracts_for_imp_lyr", "Rent_Interpolated", "med_rnt", "AVERAGE",
                             "K_NEAREST_NEIGHBORS", "", "", "", 8)

```

```{r}
# library(extrafont)
# loadfonts(device = "win")

acs_17 <- read.xlsx('Data/acs20175_tr.xlsx')
snc_tracts <- read.xlsx('Data/SNC tracts.xlsx')
housing_snc <- fread('Data/acs20175_new_builds.csv')

snc_tracts_tidy <- snc_tracts %>%
  mutate(GEOID = as.character(GEOID))

housing_snc_tidy <- housing_snc %>%
  mutate(GEOID2 = as.character(GEOID2)) %>%
  select(GEOID2, perc_new_bld)

# Remove extraneous variables
snc_acs_vars <- acs_17 %>%
  transmute(tract = GEOG_UNIT,
            pop_total = POPTOTAL,
            age65up = AGE65UP,
            white_nh = WHITENH,
            median_hhi = MEDIANHHI) %>%
  mutate(perc_poc = (pop_total - white_nh)/pop_total*100,
         perc_65up = age65up/pop_total*100) %>%
  select(-pop_total, -white_nh, -age65up)

snc_17 <- inner_join(snc_acs_vars, snc_tracts_tidy, by = c('tract' = 'GEOID')) # Select SNC tracts

snc_17_acs <- left_join(snc_17, housing_snc_tidy, by = c('tract' = 'GEOID2'))

# Bring in imputed rent
rent_imp <- read_csv('Data/acs_rent_with_imp.csv')

rent <- rent_imp %>%
  mutate(rent = ifelse(is.na(MEDGRENT), rent_imputed, MEDGRENT)) %>%
  select(GEOID, rent) %>%
  mutate(GEOID = as.character(GEOID))

# Bring in median home value
emv <- fread("Data/parcel_emv_2017.csv")

emv_tidy <- emv %>%
  mutate(Tract = as.character(Tract)) %>%
  unique()

tenures <- acs_17 %>%
  select(GEOG_UNIT, OWNEROCC, RENTEROCC) %>%
  rename(tract = GEOG_UNIT,
         owner_occ = OWNEROCC,
         rent_occ = RENTEROCC) %>%
  mutate(rent_tenure = rent_occ/(owner_occ + rent_occ),
         owner_tenure = owner_occ/(owner_occ + rent_occ)) %>%
  select(-rent_occ, -owner_occ) %>%
  unique()

# Join all data (excluding tenures, which don't get standardized) together
snc_rent <- left_join(snc_17_acs, rent, by = c("tract" = "GEOID"))
snc_full <- left_join(snc_rent, emv_tidy, by = c("tract" = "Tract"))

snc_for_cluster <- snc_full %>%
  unique()

# Create z-scores
set.seed(413)

## Results of uncapped scaling
z_no_cap <- as_tibble(scale(snc_for_cluster[-1]))
summary(z_no_cap)

## Results of capped scaling
z_cap <- replace(z_no_cap, z_no_cap > 3, 3)
z_cap <- replace(z_cap , z_cap < -3, -3)
summary(z_cap)

snc_tr_order <- snc_for_cluster %>%
  select(tract)

z_cap_tr <- bind_cols(z_cap, snc_tr_order)

z_cap_tenures <- left_join(z_cap_tr, tenures, by = "tract")

## Create composite of median home value and median rent and drop the intermediaries
z_cap_wt <- z_cap_tenures %>%
  mutate(home_value_rent_comp = (Median_EMV_2017*owner_tenure) + rent*rent_tenure) %>%
  select(-Median_EMV_2017, -owner_tenure, -rent, -rent_tenure)

z_cap_clusts <- z_cap_wt %>%
  select(-tract) # Make sure not to include tract in clustering!

# Create distance matrix of z-scores
z_matrix <- as.matrix(dist(z_cap_clusts), labels = TRUE)
fviz_dist(dist(z_cap_wt)) # Look at distance matrix at a glance

# Label rows and columns with tract name
# colnames(z_matrix) <- snc_tr_order[['tract']]
# rownames(z_matrix) <- snc_tr_order[['tract']]

fviz_nbclust(z_cap_clusts, pam, method = "silhouette")

set.seed(413)
options(scipen=999)
pam_fit_3 <- pam(z_matrix, diss = TRUE, k = 3)
plot(pam_fit_3)

snc_for_viz <- snc_full %>%
  unique() %>%
  select(-tract)

pam_fit_3$data <- z_matrix

require(cluster)
fviz_cluster(pam_fit_3, geom = "point", ellipse.type = "norm") +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  labs(title = "Clusters along Principal Components 1 & 2") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank())

clusters_17 <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_fit_3$clustering))

# clusters_17r <- clusters_17 %>%
#   mutate(color15 = ifelse(cluster==1,"Red",ifelse(cluster==2,"Blue","Gold")))

clusters_and_zdata <- left_join(clusters_17, z_cap_wt, by = "tract")

cluster_viz <- clusters_and_zdata %>%
  rename(`Percent New Builds` = perc_new_bld,
         `Percent People of Color` = perc_poc,
         `Percent 65+` = perc_65up,
         `Median Household Income` = median_hhi,
         `Home Value-Rent Composite` = home_value_rent_comp) %>%
  gather(`Percent New Builds`, `Percent People of Color`, `Percent 65+`, `Median Household Income`, `Home Value-Rent Composite`, key = "variable", value = "value")

cluster_viz %>%
  mutate(cluster = as.factor(cluster)) %>%
ggplot( aes(value, fill = cluster, color = cluster)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  facet_wrap(~variable) +
  labs(title = "Z-score Distributions by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Non-log Z-score Distributions by Cluster.png")

clusters_and_data <- left_join(clusters_17, snc_for_cluster, by = "tract")

```

# Generate plots of non-log-transformed v. log-transformed variables

```{r}
snc_17_long <- snc_full %>%
  gather(perc_poc, perc_new_bld, perc_65up, median_hhi, rent, Median_EMV_2017, key = 'Variable_category', value = 'Variable_n') %>%
  mutate(Variable_log_n = log(Variable_n)) %>%
  gather(Variable_n, Variable_log_n, key = "Transformation", value = "n") %>%
  mutate(Transformation = ifelse(Transformation == "Variable_n", "None", "Logarithmic"))

vars <- list('perc_poc', 'perc_new_bld', 'perc_65up', 'median_hhi', 'rent', 'Median_EMV_2017')

# Generate plotly plots of log-transformed v. no transformation of variables
plotly_gen <- function(filter_var) {

  filter_var <- as.character(filter_var)
  
  plot_subject <- ifelse(filter_var == 'perc_poc', '% POC',
                         ifelse(filter_var  == 'perc_new_bld', '% NEW BUILDS',
                                ifelse(filter_var  == 'perc_65up', '% 65+',
                                       ifelse(filter_var  == 'median_hhi', 'MEDIAN HHI',
                                              ifelse(filter_var == 'rent', "MEDIAN GROSS RENT",
                                                     ifelse(filter_var == 'Median_EMV_2017', "MEDIAN PARCEL EMV", 'OTHER VARIABLE'))))))

  plot <- snc_17_long %>%
  filter(Variable_category == filter_var) %>%
  ggplot(aes(n)) +
  geom_density(alpha = 0.5, fill = '#006666', color = '#006666') +
  labs(title = paste('DISTRIBUTION OF ', plot_subject, ' ACROSS METRO TRACTS'),
       x = plot_subject) +
  theme_minimal() +
  theme(text = element_text(family = "Tw Cen MT", hjust = 0.5),
        plot.title = element_text(hjust = 0.5, size = 20),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = 'white', color = 'white')) +
    facet_wrap(~Transformation)
  
ggplotly(plot)
  
}

purrr::map(vars, plotly_gen)

# See transformation v. non-transformed on their own

plotly_gen_res <- function(filter_var, transformation) {

  filter_var <- as.character(filter_var)
  
  plot_subject <- ifelse(filter_var == 'perc_poc', '% POC',
                         ifelse(filter_var  == 'perc_new_bld', '% NEW BUILDS',
                                ifelse(filter_var  == 'perc_65up', '% 65+',
                                       ifelse(filter_var  == 'median_hhi', 'MEDIAN HHI',
                                              ifelse(filter_var == 'rent', "MEDIAN GROSS RENT",
                                                     ifelse(filter_var == 'Median_EMV_2017', "MEDIAN PARCEL EMV", 'OTHER VARIABLE'))))))

  plot <- snc_17_long %>%
  filter(Variable_category == filter_var & Transformation == transformation) %>%
  ggplot(aes(n)) +
  geom_density(alpha = 0.5, fill = '#006666', color = '#006666') +
  labs(title = paste('DISTRIBUTION OF ', plot_subject, ' ACROSS METRO TRACTS'),
       x = plot_subject) +
  theme_minimal() +
  theme(text = element_text(family = "Tw Cen MT", hjust = 0.5),
        plot.title = element_text(hjust = 0.5, size = 20),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = 'white', color = 'white'))
  
ggplotly(plot)
  
}

plotly_gen_res("perc_poc", "Logarithmic")
plotly_gen_res("perc_poc", "None")

```

# Try clustering with log transformations, same seed (413)

```{r}
snc_log <- snc_for_cluster %>%
  gather(rent, Median_EMV_2017, perc_65up, perc_new_bld, perc_poc, median_hhi, key = "Category", value = "n") %>%
  mutate(n = ifelse(n != 0, log(n), 0)) %>%
  spread(Category, value = n)
  
# Create z-scores
set.seed(413)

## Results of uncapped scaling
z_no_cap_log <- as_tibble(scale(snc_log[-1]))
summary(z_no_cap_log)

## Results of capped scaling
z_cap_log <- replace(z_no_cap_log, z_no_cap_log > 3, 3)
z_cap_log <- replace(z_cap_log , z_cap_log < -3, -3)
summary(z_cap_log)

snc_tr_order <- snc_log %>%
  select(tract)

z_log_tr <- bind_cols(z_cap_log, snc_tr_order)

z_log_tenures <- left_join(z_log_tr, tenures, by = "tract")

## Create composite of median home value and median rent and drop the intermediaries
z_log_wt <- z_log_tenures %>%
  mutate(home_value_rent_comp = (Median_EMV_2017*owner_tenure) + rent*rent_tenure) %>%
  select(-Median_EMV_2017, -owner_tenure, -rent, -rent_tenure)

z_log_clusts <- z_log_wt %>%
  select(-tract) # Make sure not to include tract in clustering!

# Create distance matrix of z-scores
z_log_matrix <- as.matrix(dist(z_log_clusts), labels = TRUE)
fviz_dist(dist(z_log_wt)) # Look at distance matrix at a glance

# Label rows and columns with tract name
# colnames(z_matrix) <- snc_tr_order[['tract']]
# rownames(z_matrix) <- snc_tr_order[['tract']]

fviz_nbclust(z_log_clusts, pam, method = "silhouette")

set.seed(413)
options(scipen=999)
pam_log <- pam(z_log_matrix, diss = TRUE, k = 3)
plot(pam_log)

pam_log$data <- z_log_matrix

medoids <- as_tibble(pam_log$medoids) %>% rename(Medoid = value)
clus_width <- as_tibble(pam_log$silinfo$clus.avg.widths) %>% rename(`Average Width` = value)
pam_log_df <- bind_cols(medoids, clus_width)
av_width <- as_tibble(pam_log$silinfo$avg.width) %>%
  mutate(av_width_equals = "Average cluster width = ",
         value = round(value, 2)) %>%
  unite(label, av_width_equals, value, sep = "")

ggplot(pam_log_df, aes(Medoid, `Average Width`)) +
  geom_bar(stat = "identity", width = 0.2, fill = "#FFCC33") +
  geom_text(aes("207", 0.25), label = av_width$label, family = "Tw Cen MT") +
  labs(title = "Silhouette Width by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
        plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Cluster Silhouettes.png", width = 10, height = 6)

clusters_log <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_log$clustering))
clusters_log_sil <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_log$clustering))

clusters_and_log_z <- left_join(clusters_log, z_log_wt, by = "tract")

cluster_log_viz <- clusters_and_log_z %>%
  rename(`Percent New Builds` = perc_new_bld,
         `Percent People of Color` = perc_poc,
         `Percent 65+` = perc_65up,
         `Median Household Income` = median_hhi,
         `Home Value-Rent Composite` = home_value_rent_comp) %>%
  gather(`Percent New Builds`, `Percent People of Color`, `Percent 65+`, `Median Household Income`, `Home Value-Rent Composite`, key = "variable", value = "value")

cluster_log_viz %>%
  mutate(cluster = as.factor(cluster)) %>%
ggplot( aes(value, fill = cluster, color = cluster)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  facet_wrap(~variable) +
  labs(title = "Z-score Distributions by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Log Z-score Distributions by Cluster.png", width = 10, height = 6)

fviz_cluster(pam_log, geom = "point", ellipse.type = "norm") +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  labs(title = "Clusters along Principal Components 1 & 2") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank())

ggsave("Results/Clusters Along PCs 1 & 2.png", width = 10, height = 6)

```

Change Typology | 2000 | 201X
--------------------------------
A | 2 | 2
B | 2 | 3
C | 3 | 2 -or*-
C | 1 | 2
D | 3 | 3
E | 3 | 1 -or*-
E | 2 | 1
F | 1 | 3
G | 1 | 1

*Given that the clusters with most unlike distributions (for most variables) are 2 and 1.

```{r}
x2000 <- read.xlsx("Data/z_scores2000.xlsx")
x2015 <- read.xlsx("Data/z_scores2015.xlsx")

x2000_t <- x2000 %>%
  select(TRACT, cluster) %>%
  rename(cluster_2000 = cluster)

x2015_t <- x2015 %>%
  select(TRACT, cluster) %>%
  rename(cluster_2015 = cluster)

x2017 <- clusters_log %>%
  rename(cluster_2017 = cluster) %>%
  mutate(tract = as.numeric(tract))

change_00_15 <- left_join(x2000_t, x2015_t, by = "TRACT")
change <- left_join(change_00_15, x2017, by = c("TRACT" = "tract"))

typologies <- change %>%
  mutate(Typology_15 = ifelse(cluster_2000 == 2 & cluster_2015 == 2, "A",
                              ifelse(cluster_2000 == 2 & cluster_2015 == 3, "B",
                                     ifelse(cluster_2000 == 3 & cluster_2015 == 2 | cluster_2000 == 1 & cluster_2015 == 2, "C",
                                            ifelse(cluster_2000 == 3 & cluster_2015 == 3, "D",
                                                   ifelse(cluster_2000 == 3 & cluster_2015 == 1 | cluster_2000 == 2 & cluster_2015 == 1, "E",
                                                          ifelse(cluster_2000 == 1 & cluster_2015 == 3, "F",
                                                                 ifelse(cluster_2000 == 1 & cluster_2015 == 1, "G", "other"))))))),
         Typology_17 = ifelse(cluster_2000 == 2 & cluster_2017 == 2, "A",
                              ifelse(cluster_2000 == 2 & cluster_2017 == 3, "B",
                                     ifelse(cluster_2000 == 3 & cluster_2017 == 2 | cluster_2000 == 1 & cluster_2017 == 2, "C",
                                            ifelse(cluster_2000 == 3 & cluster_2017 == 3, "D",
                                                   ifelse(cluster_2000 == 3 & cluster_2017 == 1 | cluster_2000 == 2 & cluster_2017 == 1, "E",
                                                          ifelse(cluster_2000 == 1 & cluster_2017 == 3, "F",
                                                                 ifelse(cluster_2000 == 1 & cluster_2017 == 1, "G", "other"))))))),
         Typology_change = ifelse(Typology_15 != Typology_17, 1, 0))

typologies %>%
  group_by(Typology_change) %>%
  count()

typologies %>%
  filter(is.na(Typology_change))

write_csv(typologies, "Results/Typologies in 2000, 2015, and 2017.csv")

```

-----------------------------------

:: The above concludes work done on two different clusters, 2000 & 2017

----------------------------------

---------------------------------

:: Below begins work done on a single cluster, the change between 2000 & 2017

--------------------------------

# Read in and prep 2000 and 2017 data

```{r}
# Process 2000 data
x2000 <- read_csv("Data/cleaned_variables2000.csv") # Cleaned in previous iteration of project by Dennis Farmer
snc_tr <- read.xlsx("Data/SNC tracts.xlsx")
weight_00 <- read.xlsx("Data/NHGIS00BY10CT.xlsx")

# Get rent tenure (for weighting EMVs for parcels/median gross rent)
tenure_00 <- weight_00 %>% 
  transmute(TRACT = as.numeric(TR10),
            rent_tenure = RENTOCCW/(OWNOCCW+RENTOCCW),
            own_tenure = OWNOCCW/(OWNOCCW+RENTOCCW)) %>% 
  mutate(rent_tenure = ifelse(is.na(rent_tenure), 0, rent_tenure),
         own_tenure = ifelse(is.na(own_tenure), 0, own_tenure))

# Three tracts missing tenure data entirely
tenure_00 %>%
  filter(rent_tenure == 0 | own_tenure == 0)

snc_00 <- inner_join(x2000, snc_tr, by = c("TRACT" = "GEOID"))

snc_00_tidy <- snc_00 %>%
  select(TRACT, percentPOC2000, perSixFive2000, medHHinc2000, MedianValue2000, rent2000, builtSince00_2000) %>%
  unique()


# Process 2017 data

acs_17 <- read.xlsx('Data/acs20175_tr.xlsx')
snc_tracts <- read.xlsx('Data/SNC tracts.xlsx')
housing_snc <- fread('Data/acs20175_new_builds.csv')

snc_tracts_tidy <- snc_tracts %>%
  mutate(GEOID = as.character(GEOID))

housing_snc_tidy <- housing_snc %>%
  mutate(GEOID2 = as.character(GEOID2)) %>%
  select(GEOID2, perc_new_bld)

# Remove extraneous variables
snc_acs_vars <- acs_17 %>%
  transmute(tract = GEOG_UNIT,
            pop_total = POPTOTAL,
            age65up = AGE65UP,
            white_nh = WHITENH,
            median_hhi = MEDIANHHI) %>%
  mutate(perc_poc = (pop_total - white_nh)/pop_total*100,
         perc_65up = age65up/pop_total*100) %>%
  select(-pop_total, -white_nh, -age65up)

snc_17 <- inner_join(snc_acs_vars, snc_tracts_tidy, by = c('tract' = 'GEOID')) # Select SNC tracts

snc_17_acs <- left_join(snc_17, housing_snc_tidy, by = c('tract' = 'GEOID2'))

# Bring in imputed rent
rent_imp <- read_csv('Data/acs_rent_with_imp.csv')

rent <- rent_imp %>%
  mutate(rent = ifelse(is.na(MEDGRENT), rent_imputed, MEDGRENT)) %>%
  select(GEOID, rent) %>%
  mutate(GEOID = as.character(GEOID))

# Bring in median home value
emv <- fread("Data/parcel_emv_2017.csv")

emv_tidy <- emv %>%
  mutate(Tract = as.character(Tract)) %>%
  unique()

tenures <- acs_17 %>%
  select(GEOG_UNIT, OWNEROCC, RENTEROCC) %>%
  rename(tract = GEOG_UNIT,
         owner_occ = OWNEROCC,
         rent_occ = RENTEROCC) %>%
  mutate(rent_tenure = rent_occ/(owner_occ + rent_occ),
         owner_tenure = owner_occ/(owner_occ + rent_occ)) %>%
  select(-rent_occ, -owner_occ) %>%
  unique()

# Join all data (excluding tenures, which don't get standardized) together
snc_17_rent <- left_join(snc_17_acs, rent, by = c("tract" = "GEOID"))
snc_17_full <- left_join(snc_17_rent, emv_tidy, by = c("tract" = "Tract"))

snc_17_tidy <- snc_17_full %>%
  unique() %>%
  rename(emv = Median_EMV_2017) %>%
  mutate(tract = as.numeric(tract))

snc_00_17 <- inner_join(snc_00_tidy, snc_17_tidy, by = c("TRACT" = "tract"))

write_csv(snc_00_17, "Data/2000 and 2017 Values for SNC Tracts and Variables.csv")

```

# Cluster on the *absolute difference* between 2000 and 2017 values

```{r}
x00_17 <- read_csv("Data/2000 and 2017 Values for SNC Tracts and Variables.csv")

abs_diff <- x00_17 %>%
  mutate(poc_diff = perc_poc-percentPOC2000*100, # Convert decimal percent to whole points
         age_diff = perc_65up-perSixFive2000*100, # Convert decimal percent to whole points
         hhi_diff = median_hhi-medHHinc2000,
         emv_diff = emv-MedianValue2000,
         rent_diff = rent-rent2000,
         bld_diff = perc_new_bld-builtSince00_2000*100) %>%
  select(TRACT, poc_diff, age_diff, hhi_diff, emv_diff, rent_diff, bld_diff)

write_csv(abs_diff, "Data/Absolute Difference Between 2000 and 2017 for SNC Tracts and Variables.csv")
```

# Generate distribution plots of variables

```{r}
abs_diff_long <- abs_diff %>%
  gather(poc_diff, age_diff, hhi_diff, emv_diff, rent_diff, bld_diff, key = 'Category', value = 'n')

vars <- list('poc_diff', 'bld_diff', 'age_diff', 'hhi_diff', 'rent_diff', 'emv_diff')

# Generate plotly plots of log-transformed v. no transformation of variables
plotly_gen <- function(filter_var) {

  filter_var <- as.character(filter_var)
  
  plot_subject <- ifelse(filter_var == 'poc_diff', 'DIFFERENCE IN % POC',
                         ifelse(filter_var  == 'bld_diff', 'DIFFERENCE IN % NEW BUILDS',
                                ifelse(filter_var  == 'age_diff', 'DIFFERENCE IN % 65+',
                                       ifelse(filter_var  == 'hhi_diff', 'DIFFERENCE IN MEDIAN HHI',
                                              ifelse(filter_var == 'rent_diff', "DIFFERENCE IN MEDIAN GROSS RENT",
                                                     ifelse(filter_var == 'emv_diff', "DIFFERENCE IN MEDIAN PARCEL EMV", 'OTHER VARIABLE'))))))

  plot <- abs_diff_long %>%
  filter(Category == filter_var) %>%
  ggplot(aes(n)) +
  geom_density(alpha = 0.5, fill = '#006666', color = '#006666') +
  labs(title = paste('DISTRIBUTION OF ', plot_subject, ' ACROSS METRO TRACTS'),
       x = plot_subject) +
  theme_minimal() +
  theme(text = element_text(family = "Tw Cen MT", hjust = 0.5),
        plot.title = element_text(hjust = 0.5, size = 20),
      panel.grid = element_blank(),
      panel.background = element_rect(fill = 'white', color = 'white'))
  
ggplotly(plot)
  
}

purrr::map(vars, plotly_gen)

```

Distributions look pretty good without transformation.  Important because we can't logarithmically transform a negative number and we don't want to lose the sign of the change (+/-).  Without the sign, growth and decline will be treated synonymously.  So we'll go ahead without logarithmic transformation.

# Try clustering without log transformations, same seed (413) and experimenting with different seeds

```{r}
# Create z-scores
set.seed(413)

# Results of uncapped scaling - cannot transform logarithmically because some changes are negative & the direction is important
z_no_cap <- as_tibble(scale(abs_diff[-1]))
summary(z_no_cap)

snc_tr_order <- abs_diff %>%
  select(TRACT)

# Create distance matrix of z-scores
z_matrix <- as.matrix(dist(z_no_cap), labels = TRUE)
fviz_dist(dist(z_matrix)) # Look at distance matrix at a glance

fviz_nbclust(z_matrix, pam, method = "silhouette")

set.seed(413)
options(scipen=999)
pam_abs_diff <- pam(z_matrix, diss = TRUE, k = 3)
plot(pam_abs_diff)

set.seed(92)
options(scipen=999)
pam_abs_diff <- pam(z_matrix, diss = TRUE, k = 3)
plot(pam_abs_diff)

pam_abs_diff$data <- z_matrix

medoids <- as_tibble(pam_abs_diff$medoids) %>% rename(Medoid = value)
clus_width <- as_tibble(pam_abs_diff$silinfo$clus.avg.widths) %>% rename(`Average Width` = value)
pam_abs_diff_df <- bind_cols(medoids, clus_width)
av_width <- as_tibble(pam_abs_diff$silinfo$avg.width) %>%
  mutate(av_width_equals = "Average cluster width = ",
         value = round(value, 2)) %>%
  unite(label, av_width_equals, value, sep = "")

ggplot(pam_abs_diff_df, aes(Medoid, `Average Width`)) +
  geom_bar(stat = "identity", width = 0.2, fill = "#FFCC33") +
  geom_text(aes("361", 0.25), label = av_width$label, family = "Tw Cen MT") +
  labs(title = "Silhouette Width by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
        plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Cluster Silhouettes.png", width = 10, height = 6)

clusters_abs <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_abs_diff$clustering))
tract_z <- as_tibble(bind_cols(tract = snc_tr_order[,1], z_no_cap))
clusters_and_z <- left_join(clusters_abs, tract_z, by = "TRACT")
clusters_and_data <- left_join(clusters_abs, abs_diff, by = "TRACT")

clusters_abs_viz <- clusters_and_z %>%
  rename(`Percent New Builds` = bld_diff,
         `Percent People of Color` = poc_diff,
         `Percent 65+` = age_diff,
         `Median Household Income` = hhi_diff,
         `Median Gross Rent` = rent_diff,
         `Median Estimated Market Value` = emv_diff) %>%
  gather(`Percent New Builds`, `Percent People of Color`, `Percent 65+`, `Median Household Income`, `Median Gross Rent`, `Median Estimated Market Value`, key = "variable", value = "value")

clusters_abs_viz %>%
  mutate(cluster = as.factor(cluster)) %>%
  filter(variable == "Percent New Builds") %>%
ggplot( aes(value, fill = cluster, color = cluster)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  facet_wrap(~variable) +
  labs(title = "Z-score Distributions by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))
#ggsave("Results/Log Z-score Distributions by Cluster.png", width = 10, height = 6)

clusters_diff_viz <- clusters_and_data %>%
  rename(`Percent New Builds` = bld_diff,
         `Percent People of Color` = poc_diff,
         `Percent 65+` = age_diff,
         `Median Household Income` = hhi_diff,
         `Median Gross Rent` = rent_diff,
         `Median Estimated Market Value` = emv_diff) %>%
  gather(`Percent New Builds`, `Percent People of Color`, `Percent 65+`, `Median Household Income`, `Median Gross Rent`, `Median Estimated Market Value`, key = "variable", value = "value")

clusters_diff_viz %>%
  mutate(cluster = as.factor(cluster)) %>%
  filter(variable == "Percent New Builds") %>%
ggplot( aes(value, fill = cluster, color = cluster)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  facet_wrap(~variable) +
  labs(title = "Z-score Distributions by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

hist <- clusters_diff_viz %>%
  filter(variable == "Percent People of Color") %>%
ggplot(aes(value)) +
  geom_histogram(alpha = 0.7, fill = "#66CCCC") +
  labs(title = "Histogram of % Change") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

ggplotly(hist)
```

# Cluster on the *percent difference* between 2000 and 2017 values

```{r}
#x00_17 <- read_csv("Data/2000 and 2017 Values for SNC Tracts and Variables.csv")

perc_diff <- x00_17 %>%
  mutate(builtSince00_2000 = ifelse(builtSince00_2000 == 0, 1, builtSince00_2000)) %>%
  mutate(poc_diff = (perc_poc-percentPOC2000*100), # Convert decimal percent to whole points
         age_diff = (perc_65up-perSixFive2000*100), # Convert decimal percent to whole points
         hhi_diff = (median_hhi-medHHinc2000)/medHHinc2000*100,
         emv_diff = (emv-MedianValue2000)/MedianValue2000*100,
         rent_diff = (rent-rent2000)/rent2000*100,
         bld_diff = (perc_new_bld-builtSince00_2000*100)) %>%
  select(TRACT, poc_diff, age_diff, hhi_diff, emv_diff, rent_diff, bld_diff) %>%
  filter(!is.na(emv_diff))

write_csv(perc_diff, "Data/Percent Difference Between 2000 and 2017 for SNC Tracts and Variables.csv")
```

# Percent difference with z-scores

```{r}
# Create z-scores
set.seed(413)

# Results of uncapped scaling - cannot transform logarithmically because some changes are negative & the direction is important
z_no_cap <- as_tibble(scale(perc_diff[-1]))
summary(z_no_cap)

snc_tr_order <- perc_diff %>%
  select(TRACT)

# Create distance matrix of z-scores
z_matrix <- as.matrix(dist(z_no_cap), labels = TRUE)
fviz_dist(dist(z_matrix)) # Look at distance matrix at a glance

fviz_nbclust(z_matrix, pam, method = "silhouette")

set.seed(413)
options(scipen=999)
pam_perc_diff <- pam(z_matrix, diss = TRUE, k = 3)
plot(pam_perc_diff)

set.seed(92)
options(scipen=999)
pam_perc_diff <- pam(z_matrix, diss = TRUE, k = 3)
plot(pam_perc_diff)

pam_perc_diff$data <- z_matrix

medoids <- as_tibble(pam_perc_diff$medoids) %>% rename(Medoid = value)
clus_width <- as_tibble(pam_perc_diff$silinfo$clus.avg.widths) %>% rename(`Average Width` = value)
pam_perc_diff_df <- bind_cols(medoids, clus_width)
av_width <- as_tibble(pam_perc_diff$silinfo$avg.width) %>%
  mutate(av_width_equals = "Average cluster width = ",
         value = round(value, 2)) %>%
  unite(label, av_width_equals, value, sep = "")

ggplot(pam_perc_diff_df, aes(Medoid, `Average Width`)) +
  geom_bar(stat = "identity", width = 0.2, fill = "#FFCC33") +
  geom_text(aes("361", 0.25), label = av_width$label, family = "Tw Cen MT") +
  labs(title = "Silhouette Width by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
        plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Cluster Silhouettes.png", width = 10, height = 6)

clusters_perc <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_perc_diff$clustering))
tract_z <- as_tibble(bind_cols(tract = snc_tr_order[,1], z_no_cap))
clusters_and_z <- left_join(clusters_perc, tract_z, by = "TRACT")
clusters_and_data <- left_join(clusters_perc, perc_diff, by = "TRACT")

clusters_perc_viz <- clusters_and_z %>%
  rename(`Percent New Builds` = bld_diff,
         `Percent People of Color` = poc_diff,
         `Percent 65+` = age_diff,
         `Median Household Income` = hhi_diff,
         `Median Gross Rent` = rent_diff,
         `Median Estimated Market Value` = emv_diff) %>%
  gather(`Percent New Builds`, `Percent People of Color`, `Percent 65+`, `Median Household Income`, `Median Gross Rent`, `Median Estimated Market Value`, key = "variable", value = "value")

clusters_perc_viz %>%
  mutate(cluster = as.factor(cluster)) %>%
ggplot( aes(value, fill = cluster, color = cluster)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  facet_wrap(~variable) +
  labs(title = "Z-score Distributions by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))
#ggsave("Results/Log Z-score Distributions by Cluster.png", width = 10, height = 6)

clusters_diff_viz <- clusters_and_data %>%
  rename(`Percent New Builds` = bld_diff,
         `Percent People of Color` = poc_diff,
         `Percent 65+` = age_diff,
         `Median Household Income` = hhi_diff,
         `Median Gross Rent` = rent_diff,
         `Median Estimated Market Value` = emv_diff) %>%
  gather(`Percent New Builds`, `Percent People of Color`, `Percent 65+`, `Median Household Income`, `Median Gross Rent`, `Median Estimated Market Value`, key = "variable", value = "value")

clusters_diff_viz %>%
  mutate(cluster = as.factor(cluster)) %>%
  filter(variable == "Percent New Builds") %>%
ggplot( aes(value, fill = cluster, color = cluster)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  facet_wrap(~variable) +
  labs(title = "Z-score Distributions by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

hist <- clusters_diff_viz %>%
  filter(variable == "Percent People of Color") %>%
ggplot(aes(value)) +
  geom_histogram(alpha = 0.7, fill = "#66CCCC") +
  labs(title = "Histogram of % Change") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

ggplotly(hist)
```

# Percent difference without z-scores (because all in the same units - % - & same scale - -inf to inf)

```{r}
set.seed(413)

perc_diff_clust <- perc_diff %>% select(-TRACT)

snc_tr_order <- perc_diff %>%
  select(TRACT)

# Create distance matrix of percentages
z_matrix <- as.matrix(dist(perc_diff_clust), labels = TRUE)
fviz_dist(dist(z_matrix)) # Look at distance matrix at a glance

fviz_nbclust(z_matrix, pam, method = "silhouette")

set.seed(413)
options(scipen=999)
pam_perc_diff <- pam(z_matrix, diss = TRUE, k = 3)
plot(pam_perc_diff)

# set.seed(92)
# options(scipen=999)
# pam_perc_diff <- pam(z_matrix, diss = TRUE, k = 3)
# plot(pam_perc_diff)

pam_perc_diff$data <- z_matrix

medoids <- as_tibble(pam_perc_diff$medoids) %>% rename(Medoid = value)
clus_width <- as_tibble(pam_perc_diff$silinfo$clus.avg.widths) %>% rename(`Cluster Average Width` = value)
av_width <- as_tibble(pam_perc_diff$silinfo$avg.width)
pam_perc_diff_df <- bind_cols(medoids, clus_width) %>% mutate(`Average Width` = av_width$value)

write_csv(pam_perc_diff_df, "Results/Cluster Silhouettes - Percent Difference, 3-Cluster Solution.csv")

av_width <- as_tibble(pam_perc_diff$silinfo$avg.width) %>%
  mutate(av_width_equals = "Average cluster width = ",
         value = round(value, 2)) %>%
  unite(label, av_width_equals, value, sep = "")

ggplot(pam_perc_diff_df, aes(Medoid, `Average Width`)) +
  geom_bar(stat = "identity", width = 0.2, fill = "#FFCC33") +
  geom_text(aes("223", 0.25), label = av_width$label, family = "Tw Cen MT") +
  labs(title = "Silhouette Width by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
        plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

ggsave("Results/Cluster Silhouettes - Percent Difference, 3-Cluster Solution.png", width = 10, height = 6)

clusters_perc <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_perc_diff$clustering)) %>% mutate(cluster_type = "Percent-difference, no z-scores, 3-solution")

write_csv(clusters_perc, "Results/Percent-difference 3-Cluster Solution Clusters.csv")

tract_z <- as_tibble(bind_cols(tract = snc_tr_order[,1], perc_diff_clust))
clusters_and_z <- left_join(clusters_perc, tract_z, by = "TRACT")
clusters_and_data <- left_join(clusters_perc, perc_diff, by = "TRACT")


fviz_cluster(pam_perc_diff, geom = "point", ellipse.type = "norm") +
  scale_fill_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66", "#99CCCC")) +
  labs(title = "Clusters along Principal Components 1 & 2") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank())

ggsave("Results/Clusters Along PCs 1 & 2 - Percent Difference, 3-Cluster Solution.png", width = 10, height = 6)
```

# See a 2-cluster solution

```{r}
set.seed(413)

perc_diff_clust <- perc_diff %>% select(-TRACT)

snc_tr_order <- perc_diff %>%
  select(TRACT)

# Create distance matrix of z-scores
z_matrix <- as.matrix(dist(perc_diff_clust), labels = TRUE)

fviz_nbclust(z_matrix, pam, method = "silhouette")
ggsave("Results/Percent-Difference Average Silhouette Width by Number of Clusters.png", width = 10, height = 6)

set.seed(413)
options(scipen=999)
pam_perc_diff <- pam(z_matrix, diss = TRUE, k = 2)

pam_perc_diff$data <- z_matrix

medoids <- as_tibble(pam_perc_diff$medoids) %>% rename(Medoid = value)
clus_width <- as_tibble(pam_perc_diff$silinfo$clus.avg.widths) %>% rename(`Cluster Average Width` = value)
av_width <- as_tibble(pam_perc_diff$silinfo$avg.width)
pam_perc_diff_df <- bind_cols(medoids, clus_width) %>% mutate(`Average Width` = av_width$value)


write_csv(pam_perc_diff_df, "Results/Cluster Silhouettes - Percent-Difference, 2-Cluster Solution.csv")

av_width <- as_tibble(pam_perc_diff$silinfo$avg.width) %>%
  mutate(av_width_equals = "Average cluster width = ",
         value = round(value, 2)) %>%
  unite(label, av_width_equals, value, sep = "")

ggplot(pam_perc_diff_df, aes(Medoid, `Average Width`)) +
  geom_bar(stat = "identity", width = 0.2, fill = "#FFCC33") +
  geom_text(aes("223", 0.25), label = av_width$label, family = "Tw Cen MT") +
  labs(title = "Silhouette Width by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
        plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

ggsave("Results/Cluster Silhouettes - Percent-Difference, 2-Cluster Solution.png", width = 10, height = 6)

clusters_perc <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_perc_diff$clustering)) %>%
  mutate(cluster_type = "Percent-difference, no z-scores, 2-solution")
tract_z <- as_tibble(bind_cols(tract = snc_tr_order[,1], perc_diff_clust))
clusters_and_z <- left_join(clusters_perc, tract_z, by = "TRACT")
clusters_and_data <- left_join(clusters_perc, perc_diff, by = "TRACT")

write_csv(clusters_perc, "Results/Percent-Difference 2-Cluster Solutions Clusters.csv")

fviz_cluster(pam_perc_diff, geom = "point", ellipse.type = "norm") +
  scale_fill_manual(values = c("#FF6666", "#FFCC66")) +
  scale_color_manual(values = c("#FF6666", "#FFCC66")) +
  labs(title = "Clusters along Principal Components 1 & 2") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank())

ggsave("Results/Clusters Along PCs 1 & 2 - Percent Difference, 2 Cluster Solution.png", width = 10, height = 6)
```

# See a 7-cluster solution

```{r}
set.seed(413)
options(scipen=999)
pam_perc_diff <- pam(z_matrix, diss = TRUE, k = 7)
plot(pam_perc_diff)

pam_perc_diff$data <- z_matrix

medoids <- as_tibble(pam_perc_diff$medoids) %>% rename(Medoid = value)
clus_width <- as_tibble(pam_perc_diff$silinfo$clus.avg.widths) %>% rename(`Cluster Average Width` = value)
av_width <- as_tibble(pam_perc_diff$silinfo$avg.width)
pam_perc_diff_df <- bind_cols(medoids, clus_width) %>% mutate(`Average Width` = av_width$value)

#write_csv(pam_perc_diff_df, "Results/Cluster Silhouettes - Percent-Difference, 7-Cluster Solution.csv")

av_width <- as_tibble(pam_perc_diff$silinfo$avg.width) %>%
  mutate(av_width_equals = "Average cluster width = ",
         value = round(value, 2)) %>%
  unite(label, av_width_equals, value, sep = "")

ggplot(pam_perc_diff_df, aes(Medoid, `Average Width`)) +
  geom_bar(stat = "identity", width = 0.2, fill = "#FFCC33") +
  geom_text(aes("263", 0.3), label = av_width$label, family = "Tw Cen MT") +
  labs(title = "Silhouette Width by Cluster") +
  theme(text = element_text(family = "Tw Cen MT"),
        plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "white"))

#ggsave("Results/Cluster Silhouettes - Percent-Difference, 7-Cluster Solution.png", width = 10, height = 6)

clusters_perc <- as_tibble(bind_cols(tract = snc_tr_order[,1], cluster = pam_perc_diff$clustering)) %>%
  mutate(cluster_type = "Percent-difference, no z-scores, 7-solution")

#write_csv(clusters_perc, "Results/Percent-Difference 7-Cluster Solution Clusters.csv")

fviz_cluster(pam_perc_diff, geom = "point", ellipse.type = "norm") +
  labs(title = "Clusters along Principal Components 1 & 2") +
  theme(text = element_text(family = "Tw Cen MT"),
    plot.title = element_text(hjust = 0.5, size = 30),
        panel.background = element_rect(fill = "white"),
        legend.position = "bottom",
        axis.ticks = element_blank())

#ggsave("Results/Clusters Along PCs 1 & 2 - Percent Difference, 7-Cluster Solution.png", width = 10, height = 6)
library(stats)

pca <- stats::prcomp(pam_perc_diff$data, scale = FALSE, center = FALSE)
axes <- c(1,2)
ind <- facto_summarize(pca, element = "ind", result = "coord", axes = axes)
eig <- get_eigenvalue(pca)[axes, 2]

plot.data <- cbind.data.frame(ind, cluster = pam_perc_diff$cluster)
write_csv(plot.data, "PCA on 7 clusters.csv")

dims <- plot.data %>% select(Dim.1, Dim.2)
dims_scaled <- scale(dims)

dims_scaled_df <- as_tibble(dims_scaled)
cluster_df <- as_tibble(pam_perc_diff$clustering) %>% rename(cluster = value)
dims_cl <- bind_cols(dims_scaled_df, cluster_df)

dims_cl %>%
ggplot(aes(Dim.1*-1, Dim.2, color = cluster)) +
  geom_point() +
  geom_jitter()


```


